###========== SET UP
# LOAD LIBRARIES
```{r}
library(Seurat)
library(tidyverse)
library(patchwork)
source("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/Functions.R")
library(readr)

options(future.globals.maxSize = 100000 * 1024^10)
```

# SET UP DIRECTORIES
```{r}
setwd("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/04_CheckDbl_otherQC")
root_dir <- "/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/04_CheckDbl_otherQC"

RDSfiles <- paste0(root_dir, "/RDS_files")
dir.create(RDSfiles, showWarnings = F)

Figures_dir <- paste0(root_dir, "/Figures")
dir.create(Figures_dir, showWarnings = F)
```

# LISTS
```{r}
CanMarkers <- c(
  "SLC17A6", # glut cells,
  # "NR2E1", #adult NSCs
  # "SOX4", #neuronally committed precursors
  # "DCX", #migrating neuroblasts
  "GAD1", # gaba cells
  "PDGFRA", # OPCs
  "MBP", # Oligos
  "LRIG1", # Astro
  "CSF1R", # Microglia
  "FLT1" # Endothelial
)
```


###========= PROCESS
# LOAD ANNOTATED DATA AND 
```{r}
SeuObj <- readRDS(file = paste0(RDSfiles, "/Merged_umap_cellClass.rds"))
```


# VISUALIZE DOUBLET FLAGS
```{r}
DimPlot(SeuObj, group.by = "flag_MaybeDbl") + coord_equal()
FeaturePlot(SeuObj, features = "doublet_score") + coord_equal()

VlnPlot(SeuObj, features = "doublet_score", group.by = "flag_MaybeDbl", pt.size = 1)
DotPlot(SeuObj, features = c(CanMarkers, "doublet_score")) + RotatedAxis()
DotPlot(SeuObj, features = c(CanMarkers, "doublet_score", "BCAS1", "TCF7L2", "FOXJ1")) + RotatedAxis() + coord_flip()
```


# SUBSET BASED ON DOUBLETS FLAGS AND RE-CLUSTER
```{r}
sub_SeuObj <- subset(SeuObj, subset = flag_MaybeDbl == "NoFlag")

assayUse <- "SCT_byRegion"
reductionUse <- paste0("pca_", assayUse)

DefaultAssay(sub_SeuObj) <- "RNA"
sub_SeuObj <- DietSeurat(sub_SeuObj, assays = "RNA")
sub_SeuObj <- JoinLayers(sub_SeuObj)

# standard process
sub_SeuObj <- SCTransform(sub_SeuObj,
  new.assay.name = assayUse,
  vars.to.regress = "BirdID", return.only.var.genes = T
)
sub_SeuObj <- RunPCA(sub_SeuObj, assay = assayUse, reduction.name = reductionUse) %>%
  FindNeighbors(dims = 1:30, reduction = reductionUse) %>%
  FindClusters(resolution = 1, graph.name = paste0(assayUse, "_snn"), cluster.name = paste0("unintegrated_", reductionUse))
sub_SeuObj <- RunUMAP(sub_SeuObj,
  dims = 1:30,
  reduction = reductionUse,
  reduction.name = paste0("umap.unintegrated_", assayUse)
)

saveRDS(sub_SeuObj, file = paste0(RDSfiles, "/Merged_sctJoinOnlyVar_noDblFlagged.rds"))
```


# VISUALIZE--AND REMOVE MORE DOUBLETS
```{r}
sub_SeuObj <- readRDS(file = paste0(RDSfiles, "/Merged_sctJoinOnlyVar_noDblFlagged.rds"))

sub_SeuObj <- BuildClusterTree(sub_SeuObj, features = c(CanMarkers, "doublet_score"), reorder = T)
sub_SeuObj$Clusters2 <- Idents(sub_SeuObj)

DimPlot(sub_SeuObj, label = T) + coord_equal() + NoLegend()
ggsave(
  filename = paste0(Figures_dir, "/DblClean2/umap_clusters.tiff"),
  bg = "white", width = 7, height = 7
)

DimPlot(sub_SeuObj, group.by = "CellClass_Sct_lvl2_CxG") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/DblClean2/umap_prevClass.tiff"),
  bg = "white", width = 8, height = 7
)

FeaturePlot(sub_SeuObj, features = "doublet_score", coord.fixed = T)
ggsave(
  filename = paste0(Figures_dir, "/DblClean2/featurePlot_DblScore.tiff"),
  bg = "white", width = 7, height = 7
)

DotPlot(sub_SeuObj, features = c(CanMarkers, "doublet_score")) + RotatedAxis()
ggsave(
  filename = paste0(Figures_dir, "/DblClean2/dotPlot_CanMarkers_DblScore.tiff"),
  bg = "white", width = 6, height = 15
)
```

# EXCLUDE NEW DBL CLUSTER AND REPROCESS
Doublet clusters identified by umap location, doublet score, and can marker expression.

```{r}
sub_SeuObj_3 <- subset(sub_SeuObj, subset = Clusters2 %in% c("62", "73", "45", "66", "69"), invert = T)

DefaultAssay(sub_SeuObj_2) <- "RNA"
sub_SeuObj_2 <- DietSeurat(sub_SeuObj_2, assay = "RNA")

# standard process
sub_SeuObj_2 <- SCTransform(sub_SeuObj_2,
  new.assay.name = assayUse,
  vars.to.regress = "BirdID", return.only.var.genes = T
)
sub_SeuObj_2 <- RunPCA(sub_SeuObj_2, assay = assayUse, reduction.name = reductionUse) %>%
  FindNeighbors(dims = 1:30, reduction = reductionUse) %>%
  FindClusters(resolution = 1, graph.name = paste0(assayUse, "_snn"), cluster.name = paste0("unintegrated_", reductionUse))
sub_SeuObj_2 <- RunUMAP(sub_SeuObj_2,
  dims = 1:30,
  reduction = reductionUse,
  reduction.name = paste0("umap.unintegrated_", assayUse)
)


# Figures
sub_SeuObj_2 <- BuildClusterTree(sub_SeuObj_2, features = c(CanMarkers, "doublet_score"), reorder = T)
sub_SeuObj_2$Clusters3 <- Idents(sub_SeuObj_2)

DimPlot(sub_SeuObj_2, label = T) + coord_equal() + NoLegend()
ggsave(
  filename = paste0(Figures_dir, "/DblClean3/umap_clusters.tiff"),
  bg = "white", width = 7, height = 7
)

DimPlot(sub_SeuObj_2, group.by = "CellClass_Sct_lvl2_CxG") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/DblClean3/umap_prevClass.tiff"),
  bg = "white", width = 8, height = 7
)

FeaturePlot(sub_SeuObj_2, features = "doublet_score", coord.fixed = T)
ggsave(
  filename = paste0(Figures_dir, "/DblClean3/featurePlot_DblScore.tiff"),
  bg = "white", width = 7, height = 7
)

DotPlot(sub_SeuObj_2, features = c(CanMarkers, "doublet_score")) + RotatedAxis()
ggsave(
  filename = paste0(Figures_dir, "/DblClean3/dotPlot_CanMarkers_DblScore.tiff"),
  bg = "white", width = 6, height = 15
)

idents <- unique(Idents(sub_SeuObj_2))
plot_list <- list()

for (ident in idents) {
  p <- DimPlot(sub_SeuObj_2, cells.highlight = CellsByIdentities(sub_SeuObj_2, idents = ident)) + coord_equal()
  plot_list[[as.character(ident)]] <- p
}

combined_plot <- wrap_plots(plot_list, ncol = 8)
ggsave(
  filename = paste0(Figures_dir, "/DblClean3/umap_HiglightByClust.tiff"), plot = combined_plot,
  bg = "white", width = 40, height = 40
)


# only clusters to possible remove now
plot_list <- list()
for (ident in c("33", "55", "21", "69")) {
  p <- DimPlot(sub_SeuObj_2, cells.highlight = CellsByIdentities(sub_SeuObj_2, idents = ident)) + coord_equal()
  plot_list[[as.character(ident)]] <- p
}

combined_plot <- wrap_plots(plot_list, ncol = 2)
ggsave(
  filename = paste0(Figures_dir, "/DblClean3/umap_HiglightToRem.tiff"), plot = combined_plot,
  bg = "white", width = 12, height = 10
)

saveRDS(sub_SeuObj_2, file = paste0(RDSfiles, "/Merged_sctJoinOnlyVar_noDblFlagged_2ndClean.rds"))
```

#--- REMOVE SPREAD OUT CLUSTERS

```{r}
assayUse <- "SCT_byRegion"
reductionUse <- paste0("pca_", assayUse, "_recluster")

sub_SeuObj_3 <- subset(sub_SeuObj_2, subset = Clusters2 %in% c("55", "69"), invert = T)
sub_SeuObj_3 <- DietSeurat(sub_SeuObj_3)

# recluster
sub_SeuObj_3 <- RunPCA(sub_SeuObj_3, assay = assayUse, reduction.name = reductionUse) %>%
  FindNeighbors(dims = 1:30, reduction = reductionUse) %>%
  FindClusters(resolution = 1, graph.name = paste0(assayUse, "_snn"), cluster.name = paste0("unintegrated_", reductionUse))
sub_SeuObj_3 <- RunUMAP(sub_SeuObj_3,
  dims = 1:30,
  reduction = reductionUse,
  reduction.name = paste0("umap.unintegrated_", assayUse)
)

# saveRDS(sub_SeuObj_3, file = paste0(RDSfiles, "/Merged_sctJoinOnlyVar_noDblFlagged_3rdClean.rds"))
sub_SeuObj_3 <- readRDS(file = paste0(RDSfiles, "/Merged_sctJoinOnlyVar_noDblFlagged_3rdClean.rds"))
sub_SeuObj_3 <- BuildClusterTree(sub_SeuObj_3, features = c(CanMarkers, "doublet_score"), reorder = T)
sub_SeuObj_3$Clusters4 <- Idents(sub_SeuObj_3)

# Figures
DimPlot(sub_SeuObj_3, label = T) + coord_equal() + NoLegend()
ggsave(
  filename = paste0(Figures_dir, "/DblClean4/umap_clusters.tiff"),
  bg = "white", width = 7, height = 7
)

DimPlot(sub_SeuObj_3, group.by = "CellClass_Sct_lvl2_CxG") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/DblClean4/umap_prevClass.tiff"),
  bg = "white", width = 8, height = 7
)

DimPlot(sub_SeuObj_3, group.by = "RegionAcronym") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/DblClean4/umap_Region.tiff"),
  bg = "white", width = 8, height = 7
)

FeaturePlot(sub_SeuObj_3, features = "doublet_score", coord.fixed = T)
ggsave(
  filename = paste0(Figures_dir, "/DblClean4/featurePlot_DblScore.tiff"),
  bg = "white", width = 7, height = 7
)

DotPlot(sub_SeuObj_3, features = c(CanMarkers, "doublet_score")) + RotatedAxis()
ggsave(
  filename = paste0(Figures_dir, "/DblClean4/dotPlot_CanMarkers_DblScore.tiff"),
  bg = "white", width = 6, height = 15
)

idents <- unique(Idents(sub_SeuObj_3))
plot_list <- list()
for (ident in idents) {
  p <- DimPlot(sub_SeuObj_3, cells.highlight = CellsByIdentities(sub_SeuObj_3, idents = ident)) + coord_equal()
  plot_list[[as.character(ident)]] <- p
}

combined_plot <- wrap_plots(plot_list, ncol = 8)
ggsave(
  filename = paste0(Figures_dir, "/DblClean4/umap_HiglightByClust.tiff"), plot = combined_plot,
  bg = "white", width = 40, height = 40
)


IDs <- unique(sub_SeuObj_3$orig.ident)
Idents(sub_SeuObj_3) <- "orig.ident"
plot_list <- list()
for (id in IDs) {
  p <- DimPlot(sub_SeuObj_3, cells.highlight = CellsByIdentities(sub_SeuObj_3, idents = id), order = id) + coord_equal()
  plot_list[[as.character(id)]] <- p
}

combined_plot <- wrap_plots(plot_list, ncol = 10)
ggsave(
  filename = paste0(Figures_dir, "/DblClean4/umap_HiglightByOrigIdent.tiff"), plot = combined_plot,
  bg = "white", width = 45, height = 20, dpi = 600
)
```

###----- HBV_pub3 outlier

# HARMONY
```{r}
FeaturePlot(sub_SeuObj_3, features = c("nCount_SCT", "nFeature_SCT")) + coord_equal()

IntObj_2 <- IntegrateLayers(
  object = sub_SeuObj_3, method = HarmonyIntegration,
  orig.reduction = "pca_SCT_byRegion_recluster", new.reduction = "harmony",
  normalization.method = "SCT",
  group.by.vars = "BirdID",
  verbose = TRUE
)


IntObj <- FindNeighbors(IntObj, dims = 1:30, reduction = "harmony", graph.name = c("harmony_nn", "harmony_snn")) %>%
  FindClusters(resolution = 1, graph.name = "harmony_snn", cluster.name = "harmony_clusters") %>%
  RunUMAP(dims = 1:30, reduction = "harmony", reduction.name = "umap.harmony")


DimPlot(IntObj, group.by = "RegionAcronym") + coord_equal()
DimPlot(IntObj, reduction = "umap.harmony", group.by = "RegionAcronym") + coord_equal()
```

Doesn't seem to higher Feature or Count...? 
Let's see if integration gets rid of it,
 
Harmony integration did nothing so then will go with RPCA, see if that changes things?


####--- try two different splits

0) Merge and sct regress BirdID did not work 
1) by Region w/regress BirdID per region
2) by ID
# LOAD DATA
```{r}
sub_SeuObj_3 <- readRDS(file = paste0(RDSfiles, "/Merged_sctJoinOnlyVar_noDblFlagged_3rdClean.rds"))
DefaultAssay(sub_SeuObj_3) <- "RNA"
sub_SeuObj_byRegion <- DietSeurat(sub_SeuObj_3, assays = "RNA")
```

# SPLIT BY REGION W/BIRD ID REGRESS
also excluding cluster 57, seems like OPC doublets
```{r}
# PREP OBJECT
sub_SeuObj_byRegion[["RNA"]] <- split(sub_SeuObj_byRegion[["RNA"]], f = sub_SeuObj_byRegion$RegionAcronym)
sub_SeuObj_byRegion$Clusters4 <- Idents(sub_SeuObj_byRegion)
sub_SeuObj_byRegion <- subset(sub_SeuObj_byRegion, subset = Clusters4 == "57", invert = T)

# STD PROCESSING
assayUse <- "SCT_byRegion_rgrsID"
reductionUse <- paste0("pca_", assayUse)

sub_SeuObj_byRegion <- SCTransform(sub_SeuObj_byRegion,
  new.assay.name = assayUse,
  vars.to.regress = "BirdID", return.only.var.genes = T
)

sub_SeuObj_byRegion <- RunPCA(sub_SeuObj_byRegion, assay = assayUse, reduction.name = reductionUse)
sub_SeuObj_byRegion <- RunUMAP(sub_SeuObj_byRegion, dims = 1:30, reduction = reductionUse, reduction.name = paste0("umap.unintegrated_", reductionUse))

DimPlot(sub_SeuObj_byRegion, reduction = paste0("umap.unintegrated_", reductionUse), group.by = "RegionAcronym") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/DblClean5_chckHVCpub3/umap_ByRegion.tiff"),
  bg = "white", width = 9, height = 8
)

DimPlot(sub_SeuObj_byRegion, reduction = paste0("umap.unintegrated_", reductionUse), cols = custom_colors, group.by = "orig.ident") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/DblClean5_chckHVCpub3/umap_ByOrigIdent.tiff"),
  bg = "white", width = 10, height = 8
)

DimPlot(sub_SeuObj_byRegion, reduction = paste0("umap.unintegrated_", reductionUse), group.by = "CellClass_Sct_lvl2_CxG") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/DblClean5_chckHVCpub3/umap_ByCxGclass.tiff"),
  bg = "white", width = 9, height = 8
)

IDs <- unique(sub_SeuObj_byRegion$orig.ident)
Idents(sub_SeuObj_byRegion) <- "orig.ident"
plot_list <- list()
for (id in IDs) {
  p <- DimPlot(sub_SeuObj_byRegion, cells.highlight = CellsByIdentities(sub_SeuObj_byRegion, idents = id), order = id) + coord_equal()
  plot_list[[as.character(id)]] <- p
}
combined_plot <- wrap_plots(plot_list, ncol = 10)
ggsave(
  filename = paste0(Figures_dir, "/DblClean5_chckHVCpub3/umap_HiglightByOrigIdent.tiff"), plot = combined_plot,
  bg = "white", width = 45, height = 20, dpi = 600
)

#--- Integrate Harmony
sub_SeuObj_byRegion <- IntegrateLayers(
  object = sub_SeuObj_byRegion,
  method = HarmonyIntegration,
  orig.reduction = reductionUse,
  new.reduction = "harmony",
  normalization.method = "SCT",
  verbose = FALSE
)

sub_SeuObj_byRegion[["RNA"]] <- JoinLayers(sub_SeuObj_byRegion[["RNA"]])

sub_SeuObj_byRegion <- RunUMAP(sub_SeuObj_byRegion, reduction = "harmony", dims = 1:30, reduction.name = "umap.harmony")
# saveRDS(sub_SeuObj_byRegion, paste0(RDSfiles, "/IntegrateHarmony_20240819.rds"))
sub_SeuObj_byRegion <- readRDS(paste0(RDSfiles, "/IntegrateHarmony_20240819.rds"))

# Harmony Figures
sub_Fig_dir <- "/DblClean5_chckHVCpub3/Harmony/"

DimPlot(sub_SeuObj_byRegion, reduction = "umap.harmony", group.by = "RegionAcronym") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, sub_Fig_dir, "umap_ByRegion.tiff"),
  bg = "white", width = 9, height = 8
)

DimPlot(sub_SeuObj_byRegion, reduction = "umap.harmony", cols = custom_colors, group.by = "orig.ident") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, sub_Fig_dir, "umap_ByOrigIdent.tiff"),
  bg = "white", width = 10, height = 8
)

DimPlot(sub_SeuObj_byRegion, reduction = "umap.harmony", group.by = "CellClass_Sct_lvl2_CxG") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, sub_Fig_dir, "umap_ByCxGclass.tiff"),
  bg = "white", width = 9, height = 8
)

IDs <- unique(sub_SeuObj_byRegion$orig.ident)
Idents(sub_SeuObj_byRegion) <- "orig.ident"
plot_list <- list()
for (id in IDs) {
  p <- DimPlot(sub_SeuObj_byRegion, cells.highlight = CellsByIdentities(sub_SeuObj_byRegion, idents = id), order = id) + coord_equal()
  plot_list[[as.character(id)]] <- p
}
combined_plot <- wrap_plots(plot_list, ncol = 10)
ggsave(
  filename = paste0(Figures_dir, sub_Fig_dir, "umap_HiglightByOrigIdent.tiff"), plot = combined_plot,
  bg = "white", width = 45, height = 20, dpi = 600
)
```

### one last try to integrate HVC_Pub3
# RPCA integration
```{r}
sub_SeuObj_byRegion <- readRDS(paste0(RDSfiles, "/IntegrateHarmony_20240819.rds"))
sub_SeuObj_byRegion[["RNA"]] <- split(sub_SeuObj_byRegion[["RNA"]], f = sub_SeuObj_byRegion$RegionAcronym)

sub_SeuObj_byRegion <- IntegrateLayers(
  object = sub_SeuObj_byRegion,
  method = RPCAIntegration,
  orig.reduction = "pca_SCT_byRegion_rgrsID",
  new.reduction = "rpca.integrated",
  normalization.method = "SCT",
  verbose = FALSE
)

sub_SeuObj_byRegion[["RNA"]] <- JoinLayers(sub_SeuObj_byRegion[["RNA"]])

sub_SeuObj_byRegion <- RunUMAP(sub_SeuObj_byRegion, reduction = "rpca.integrated", dims = 1:30, reduction.name = "umap.rpca")

saveRDS(sub_SeuObj_byRegion, file = paste0(RDSfiles, "/IntegratedRPCA_20240820.rds"))

# figures
sub_Fig_dir <- "/DblClean5_chckHVCpub3/RPCA/"

DimPlot(sub_SeuObj_byRegion, reduction = "umap.rpca", group.by = "RegionAcronym") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, sub_Fig_dir, "umap_ByRegion.tiff"),
  bg = "white", width = 9, height = 8
)

DimPlot(sub_SeuObj_byRegion, reduction = "umap.rpca", cols = custom_colors, group.by = "orig.ident") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, sub_Fig_dir, "umap_ByOrigIdent.tiff"),
  bg = "white", width = 10, height = 8
)

DimPlot(sub_SeuObj_byRegion, reduction = "umap.rpca", group.by = "CellClass_Sct_lvl2_CxG") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, sub_Fig_dir, "umap_ByCxGclass.tiff"),
  bg = "white", width = 9, height = 8
)

IDs <- unique(sub_SeuObj_byRegion$orig.ident)
Idents(sub_SeuObj_byRegion) <- "orig.ident"
plot_list <- list()
for (id in IDs) {
  p <- DimPlot(sub_SeuObj_byRegion, reduction = "umap.rpca", cells.highlight = CellsByIdentities(sub_SeuObj_byRegion, idents = id), order = id) + coord_equal()
  plot_list[[as.character(id)]] <- p
}
combined_plot <- wrap_plots(plot_list, ncol = 10)
ggsave(
  filename = paste0(Figures_dir, sub_Fig_dir, "umap_HiglightByOrigIdent.tiff"), plot = combined_plot,
  bg = "white", width = 45, height = 20, dpi = 600
)
```
