---
title: "Doublet removal and other possible QC"
output: html_notebook
---

Checking if I should remove doublets from scDblFinder.
Overall it looks like scDblFinder is better than Doublet Finder, 
  with only missing a few possible doublets.

###========== SET UP
# LOAD LIBRARIES
```{r}
library(Seurat)
library(tidyverse)
source("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/Functions.R")
library(readr)
```

# SET UP DIRECTORIES
```{r}
setwd("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/04_CheckDbl_otherQC")
root_dir <- "/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/04_CheckDbl_otherQC"

RDSfiles <- paste0(root_dir, "/RDS_files")
dir.create(RDSfiles, showWarnings = F)

Figures_dir <- paste0(root_dir, "/Figures")
dir.create(Figures_dir, showWarnings = F)
```

# LISTS
```{r}
CanMarkers <- c(
  "SLC17A6", # glut cells,
  # "NR2E1", #adult NSCs
  # "SOX4", #neuronally committed precursors
  # "DCX", #migrating neuroblasts
  "GAD1", # gaba cells
  "PDGFRA", # OPCs
  "MBP", # Oligos
  "LRIG1", # Astro
  "CSF1R", # Microglia
  "FLT1" # Endothelial
)
```


###========== MERGING

# LOAD MERGED OBJECT
```{r}
Obj_merged <- readRDS(file = "/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/03_Merge_Pallium/RDS_files/Merged_SCTbyOrigIdent_umap.rds")
```

###===== REMOVE DOUBLETS?

# SHOWING DOUBLETS
```{r}
Obj_merged <- BuildClusterTree(Obj_merged, features = c(CanMarkers, "doublet_score"), reorder = T)
Obj_merged[["unintegrated_sctByOrig"]] <- Idents(object = Obj_merged)

DotPlot(Obj_merged, features = c(CanMarkers, "doublet_score")) + RotatedAxis()
ggsave(
  filename = sprintf("%s/DotPlot_unintegrated_sctByOrig_%s.tiff", Figures_dir, "CanMarkers"),
  bg = "white", width = 5, height = 12
)

groupings <- c("BirdID", "doublet_class", "unintegrated_sctByOrig")
for (groupBy in groupings) {
  DimPlot(Obj_merged, reduction = "umap.unintegrated_sctByOrig", group.by = groupBy) + coord_fixed()
  ggsave(
    filename = sprintf("%s/umap_unintegrated_sctByOrig_groupBy_%s.tiff", Figures_dir, groupBy),
    bg = "white", width = 8, height = 7
  )
}

FeaturePlot(Obj_merged, reduction = "umap.unintegrated_sctByOrig", features = "doublet_score") + coord_fixed()
ggsave(
  filename = sprintf("%s/FeatPlot_unintegrated_sctByOrig_groupBy_%s.tiff", Figures_dir, "doublet_score"),
  bg = "white", width = 8, height = 7
)

# stacked plots
StacksToPlot <- c("orig.ident", "doublet_class", "RegionAcronym")
MetaData <- Obj_merged[[c("unintegrated_sctByOrig", StacksToPlot)]]
for (StackGroup in StacksToPlot) {
  stackedbarplot(MetaData, "unintegrated_sctByOrig", StackGroup)
  ggsave(
    filename = sprintf("%s/StackPlot_unintegrated_sctByOrig_%s.tiff", Figures_dir, StackGroup),
    bg = "white", width = 10, height = 20
  )
}
```

# TESTING WAYS TO REMOVE DOUBLETS
```{r}
# Inspect potential doublets in clusters
plot <- DimPlot(Obj_merged, reduction = "umap.unintegrated_sctByOrig", group.by = "doublet_class") + coord_fixed()
# HoverLocator(plot = plot, information = FetchData(Obj_merged, vars = c("unintegrated_sctByOrig")))

DimPlot(Obj_merged, reduction = "umap.unintegrated_sctByOrig", cells = WhichCells(Obj_merged, idents = 61)) + coord_fixed()

FeaturePlot(Obj_merged,
  reduction = "umap.unintegrated_sctByOrig", features = c("CSF1R", "SLC17A6"), blend = TRUE,
  cells = WhichCells(Obj_merged, idents = 37), split.by = "doublet_class", coord.fixed = 1
)
ggsave(
  filename = sprintf("%s/FeatPlot_test_%s.tiff", Figures_dir, "blend"),
  bg = "white", width = 20, height = 20
)


FeatureScatter(Obj_merged,
  group.by = "doublet_class",
  cells = WhichCells(Obj_merged, idents = 64),
  feature1 = "CSF1R", feature2 = "LRIG1", jitter = T, shuffle = T
)
ggsave(
  filename = sprintf("%s/ScatPlot_test_%s.tiff", Figures_dir, "60"),
  bg = "white", width = 10, height = 8
)


# do doublets express more mix of genes
Doublets_only <- subset(Obj_merged, subset = doublet_class == "doublet")
DotPlot(Doublets_only, features = CanMarkers) + RotatedAxis()
ggsave(
  filename = sprintf("%s/DotPlot_unintegrated_sctByOrig_%s_doubletsOnly.tiff", Figures_dir, "CanMarkers"),
  bg = "white", width = 5, height = 12
)

singlets_only <- subset(Obj_merged, subset = doublet_class == "singlet")
DotPlot(singlets_only, features = CanMarkers) + RotatedAxis()
ggsave(
  filename = sprintf("%s/DotPlot_unintegrated_sctByOrig_%s_singletsOnly.tiff", Figures_dir, "CanMarkers"),
  bg = "white", width = 5, height = 12
)


FeatureScatter(Obj_merged,
  group.by = "doublet_class",
  # cells = WhichCells(Obj_merged, idents = 64),
  feature1 = "CSF1R", feature2 = "LRIG1", jitter = T, shuffle = T
)


# split umaps
DimPlot(Obj_merged,
  split.by = "unintegrated_sctByOrig", group.by = "doublet_class",
  ncol = 10
)
ggsave(
  filename = sprintf("%s/umap_unintegrated_sctByOrig_splitBy_%s.tiff", Figures_dir, "Cluster"),
  bg = "white", width = 30, height = 30
)
```

# This method seems to be best to show dblFinder is good for doublet removal
```{r}
# Combination of markers feature plots
GenePairsList <- combn(CanMarkers, 2, simplify = FALSE)

for (i in seq_along(GenePairsList)) {
  FeaturePlot(Obj_merged,
    features = GenePairsList[[i]], blend = T,
    max.cutoff = "q70", blend.threshold = 0.001, coord.fixed = 1
  )
  ggsave(
    filename = sprintf("%s/BlendFeatPlots/FeatPlot_blend_%sand%s.tiff", Figures_dir, GenePairsList[[i]][1], GenePairsList[[i]][2]),
    bg = "white", width = 15, height = 10
  )
}
```



###===== EXCLUDE DOUBLETS AND OTHER BAD CELLS AND RECLUSTER TO CHECK IF CLEANER
Will add back cells from cluster 57 which might be transition OPC to Oligo, when subclustering

# load data if needed
```{r}
Obj_merged <- readRDS(file = "/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/03_Merge_Pallium/RDS_files/Merged_SCTbyOrigIdent_umap.rds")
```

# subset and remove excess metadata
```{r}
Filtered_obj_merged <- subset(Obj_merged, subset = doublet_class == "singlet")

DefaultAssay(Filtered_obj_merged) <- "RNA"
SeuObj <- DietSeurat(Filtered_obj_merged, assays = "RNA")

# REMOVE EXCESS METADATA
columns_to_remove <-
  names(SeuObj@meta.data[grepl("seurat_clusters|nFeature_SCT_byOrigIdent|nCount_SCT_byOrigIdent", names(SeuObj@meta.data))])

for (i in columns_to_remove) {
  SeuObj@meta.data[[i]] <- NULL
}
```

# Recluster and check marker expression
Now I will recluster to try get rid of bird effect.
So will treat region as treatment, 
  and therefore regress out BirdID replicate per region
```{r}
SeuObj_og <- SeuObj
SeuObj <- JoinLayers(SeuObj)
SeuObj[["RNA"]] <- split(SeuObj[["RNA"]], f = SeuObj$RegionAcronym)

SeuObj <- SCTransform(SeuObj,
  new.assay.name = "SCT_byRegion",
  vars.to.regress = "BirdID", return.only.var.genes = F
)
saveRDS(SeuObj, file = paste0(RDSfiles, "/Merged_noDbl_sctByRegion.rds"))
# SeuObj <- readRDS(file = paste0(RDSfiles, "/Merged_noDbl_sctByRegion.rds"))

assayUse <- "SCT_byRegion"
reductionUse <- paste0("pca_", assayUse)
SeuObj <- RunPCA(SeuObj, assay = assayUse, reduction.name = reductionUse) %>%
  FindNeighbors(dims = 1:30, reduction = reductionUse) %>%
  FindClusters(resolution = 1, graph.name = paste0(assayUse, "_snn"), cluster.name = paste0("unintegrated_", reductionUse))
SeuObj <- RunUMAP(SeuObj, dims = 1:30, reduction = reductionUse, reduction.name = paste0("umap.unintegrated_", assayUse))

# saveRDS(SeuObj, file = paste0(RDSfiles, "/Merged_noDbl_sctByRegion_umap.rds"))
SeuObj <- readRDS(file = paste0(RDSfiles, "/Merged_noDbl_sctByRegion_umap.rds"))
```

# CAN MARKERS
```{r}
SeuObj <- BuildClusterTree(SeuObj, features = c(CanMarkers, "doublet_score"), reorder = T)
SeuObj[[paste0("unintegrated_", reductionUse)]] <- Idents(object = SeuObj)

DotPlot(SeuObj, features = c(CanMarkers, "doublet_score")) + RotatedAxis()
ggsave(
  filename = paste0(Figures_dir, "/noDoublets/DotPlot_canMarkers_noDbl.tiff"),
  bg = "white", height = 15, width = 6
)

DimPlot(SeuObj) + coord_equal()
FeaturePlot(SeuObj, features = c("PDGFRA", "MBP"), coord.fixed = T)
```
It looks like there some remnant doublets still, so will check how to remove 

# CHECK IF NEEDS INTEGRATION
```{r}
DimPlot(SeuObj,
  reduction = paste0("umap.unintegrated_", assayUse),
  group.by = "BirdID"
) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/noDoublets/umap_BirdID_noDbl.tiff"),
  bg = "white", height = 5, width = 6
)
```
Doesn't look like it's mxing well by BirdID, especially HVCpool3. 
However, as shown below, the integration seems to mix possible doublets.
So will classify cell types then subcluster
The final umap will be made at the end


# CLASSIFY CELLS FOR SUBCLUSTERING
Based on dot plot of marker and doublet score, 
  classifying possible doublets
  
```{r}
# doublet clustering flags
SeuObj$flag_MaybeDbl <- ifelse(SeuObj$unintegrated_pca_SCT_byRegion %in% c("64", "63", "72", "73", "69", "71", "61", "74", "70"), "Flag", "NoFlag")
DimPlot(SeuObj,
  reduction = paste0("umap.unintegrated_", assayUse),
  group.by = "flag_MaybeDbl"
) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/noDoublets/umap_DlbFlag_noDbl.tiff"),
  bg = "white", height = 5, width = 6
)

# cell class
SeuObj$CellClass_Sct_lvl2 <- "Empty"
cluster_class_mapping <- list(
  Glutamatergic = c(
    "25", "6", "45", "27", "34", "44", "16", "28", "57", "78", "18",
    "51", "35", "50", "38", "66", "36", "53", "33", "40", "7", "43",
    "1", "22", "4", "5", "17", "76",
    "62", "58", "41", "56", "42", "8", "20",
    "15", "77", "13"
  ),
  GABAergic = c(
    "59", "49", "2", "23", "21", "46", "47", "11", "37", "52", "48",
    "10", "54", "32", "55",
    "60"
  ),
  OPC_Oligo = c(
    "19", "31", # opc
    "61", "74", "12", "68", "70", "9", "79"
  ), # oligo
  Astro_Endo = c(
    "0", "65", "14", "29", "73", "69", "3", "30", "26", "71", "67",
    "24", "75", "64", "63", "72"
  ), # endo
  Microglia = c("39")
)

for (class_name in names(cluster_class_mapping)) {
  cluster_names <- cluster_class_mapping[[class_name]]
  cluster_cells <- SeuObj@active.ident %in% cluster_names
  SeuObj$CellClass_Sct_lvl2[cluster_cells] <- class_name
}

# saveRDS(SeuObj, file = paste0(RDSfiles, "/Merged_umap_cellClass.rds"))

# visualize
DimPlot(SeuObj,
  reduction = paste0("umap.unintegrated_", assayUse),
  group.by = c("CellClass_Sct_lvl2")
) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/noDoublets/umap_cellClass.tiff"),
  bg = "white", width = 6, height = 5
)

DimPlot(SeuObj,
  reduction = paste0("umap.unintegrated_", assayUse),
  group.by = c("CellClass_Sct_lvl2", "flag_MaybeDbl")
) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/noDoublets/umap_cellClass_DblFlag.tiff"),
  bg = "white", width = 15, height = 10
)

MetaData <- SeuObj[[c("CellClass_Sct_lvl2", "BirdID", "orig.ident")]]
stackedbarplot(MetaData, "CellClass_Sct_lvl2", "BirdID")
stackedbarplot_colors(MetaData, "CellClass_Sct_lvl2", "orig.ident")
```
The cell labeling doesn't seem to be perfect in UMAP space.
Therefore will move it to astrocyte CellxGene to hand label the rest of cells so for when
I subcluster.

# Add CellXGene labeling
```{r}
SeuObj <- readRDS(file = paste0(RDSfiles, "/Merged_umap_cellClass.rds"))

CxG_cellLabels <- read_csv(paste0(root_dir, "/CellXGene/PallMerge-cell-labels-FE3KETAG.csv"),
  skip = 2
) %>% as.data.frame()
rownames(CxG_cellLabels) <- CxG_cellLabels$index
CxG_cellLabels <- CxG_cellLabels %>% select(-index)

SeuObj <- AddMetaData(SeuObj, CxG_cellLabels)

metadata <- SeuObj[[c("CellClass_Sct_lvl2", "CellClass_Sct_lvl2_CxG")]]
metadata2 <- metadata %>%
  mutate(CellClass_Sct_lvl2_CxG = ifelse(CellClass_Sct_lvl2_CxG == "unassigned", CellClass_Sct_lvl2, CellClass_Sct_lvl2_CxG)) %>%
  select(-CellClass_Sct_lvl2)

SeuObj <- AddMetaData(SeuObj, metadata2)

DimPlot(SeuObj, group.by = "CellClass_Sct_lvl2_CxG") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/noDoublets/umap_cellClass_CxG.tiff"),
  bg = "white", width = 6, height = 5
)

saveRDS(SeuObj, file = paste0(RDSfiles, "/Merged_umap_cellClass.rds"))
```
