---
title: "Pallium Spatial_after Lei's cleaned clustering"
output: html_notebook
---

This is after I have selected the slides that are high QC,
and added metadata of slide information (in archives). I then sent to Lei,
who worked on getting clustering more informed by spatial info, 
so that there weren't one spatial dot from BG in pallium for e.g.

###===== SET UP
# LOAD LIBRARIES
```{r}
library(Seurat)
library(tidyverse)
library(readxl)
library(Matrix)
library(scrattch.hicat)

options(future.globals.maxSize = 100000 * 1024^10)
```

# SET UP DIRECTORIES
```{r}
setwd("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_Spatial")
root_dir <- "/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_Spatial"

RDSfiles <- paste0(root_dir, "/RDS_files")
dir.create(RDSfiles, showWarnings = F)

Figures_dir <- paste0(root_dir, "/Figures")
dir.create(Figures_dir, showWarnings = F)

Tables_dir <- paste0(root_dir, "/Tables")
dir.create(Tables_dir, showWarnings = F)
```

# LOAD DATA
```{r}
spatial_all <-
  readRDS(
    file = "/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_Spatial/RDS_files/FromLei_subSpatial_20240916_v4.1.rds"
  )

spatial_all$LXident <- Idents(spatial_all)
```

###===== GET PALLIUM CLUSTERS

# VISUALIZE ALL OF TELE
```{r}
# visualize to only acquire pallium clusters
SpatialDimPlot(spatial_all, group.by = "LXident", ncol = 2, pt.size.factor = 150, image.alpha = 0.2)
ggsave(
  filename = paste0(Figures_dir, "/01_AllTele_identifyPall/allTele_LXclusters.tiff"),
  bg = "white", width = 10, height = 15
)

SpatialDimPlot(spatial_all, ncol = 2, pt.size.factor = 0, image.alpha = 1)
ggsave(
  filename = paste0(Figures_dir, "/tissue.tiff"),
  bg = "white", width = 10, height = 15
)

for (i in levels(spatial_all@active.ident)) {
  print(i)
  SpatialDimPlot(spatial_all,
    cells.highlight = WhichCells(spatial_all, idents = i),
    ncol = 2, pt.size.factor = 150, image.alpha = 0.2
  )
  ggsave(
    filename = sprintf("%s/01_AllTele_identifyPall/allTele_LXcluster_%s.tiff", Figures_dir, i),
    bg = "white", width = 10, height = 15
  )
}


custom_colors <- c(
  "0" = "#ADD8E6",
  "1" = "#B0B0B0",
  "2" = "#808080",
  "3" = "#87CEEB",
  "4" = "#6495ED",
  "5" = "#4682B4",
  "6" = "#4169E1",
  "7" = "black"
)

SpatialDimPlot(spatial_all,
  group.by = "LXident", ncol = 2, pt.size.factor = 150, image.alpha = 0.2,
  cols = custom_colors
)
ggsave(
  filename = paste0(Figures_dir, "/01_AllTele_identifyPall/allTele_LXclusters_PallCols.tiff"),
  bg = "white", width = 10, height = 15
)
```

# SUBSET FOR PALLIUM
```{r}
spatial_pall <- subset(spatial_all, subset = LXident %in%
  c("1", "2", "7"), invert = T)

# Exclude by hand few cells not in pallium
spatial_coords <- spatial_pall@images[["slice11"]]@coordinates
beads_pick_11 <- rownames(spatial_coords)[spatial_coords$row > 50 & spatial_coords$col > 20]
SpatialDimPlot(spatial_pall,
  images = "slice11", cells.highlight = beads_pick_11,
  ncol = 2, pt.size.factor = 150, image.alpha = 0.2
)

spatial_coords <- spatial_pall@images[["slice12"]]@coordinates
beads_pick_12 <- rownames(spatial_coords)[spatial_coords$row > 50 & spatial_coords$col > 40]
SpatialDimPlot(spatial_pall,
  images = "slice12", cells.highlight = beads_pick_12,
  ncol = 2, pt.size.factor = 150, image.alpha = 0.2
)

spatial_coords <- spatial_pall@images[["slice21"]]@coordinates
beads_pick_21 <- rownames(spatial_coords)[spatial_coords$row > 70 & spatial_coords$col > 40]
SpatialDimPlot(spatial_pall,
  images = "slice21", cells.highlight = beads_pick_21,
  ncol = 2, pt.size.factor = 150, image.alpha = 0.2
)

excl_beads <- c(beads_pick_11, beads_pick_12, beads_pick_21)

spatial_pall <- subset(spatial_pall,
  cells = excl_beads,
  invert = T
)

spatial_pall$palliumBeads <- "pall"
beadsInPall <- FetchData(spatial_pall, vars = "palliumBeads")
saveRDS(beadsInPall, file = paste0(Tables_dir, "/PallBeads_meta_usingLXclustersNmanual.rds"))

#- for checking
SpatialDimPlot(spatial_pall,
  ncol = 2, pt.size.factor = 150, image.alpha = 0.2
)
ggsave(
  filename = paste0(Figures_dir, "/04_Pall/pallClean.tiff"),
  bg = "white", width = 10, height = 15
)

SpatialFeaturePlot(spatial_pall,
  features = "PVALB",
  ncol = 2, pt.size.factor = 150, image.alpha = 0.2
)
ggsave(
  filename = paste0(Figures_dir, "/04_Pall/pall_PVALB.tiff"),
  bg = "white", width = 10, height = 15
)
```
Lei's smoothed clustering, helps the most to select the beads to be found in pallium.
The clusters themselves don't seem the best. I did other analysis in this script but mostly wanted just to get the beads in Pallium info.

Will do integration of just pallium.
This will give me the clusters for pallium,
    and non_pall can keep the past identity.
This is because some clusters were still spilling over

###===== RE-NORMALIZE AND INTEGRATE DATA JUST THE PALLIUM
```{r}
spatial_pall[["Spatial"]] <- split(spatial_pall[["Spatial"]], f = spatial_pall$Slice)
spatial_pall <- SCTransform(spatial_pall,
  assay = "Spatial",
  vars.to.regress = c("nCount_Spatial", "nFeature_Spatial"),
  return.only.var.genes = F,
  new.assay.name = "SCTbySlice_pall"
)

reductionUse <- "pca_sctSlice_pall"
spatial_pall <- RunPCA(spatial_pall, assay = "SCTbySlice_pall", reduction.name = reductionUse)

reductionInt <- paste0("integrated.rpca_", reductionUse)
spatial_pall <- IntegrateLayers(
  object = spatial_pall, method = RPCAIntegration,
  orig.reduction = reductionUse, new.reduction = reductionInt,
  normalization.method = "SCT", verbose = FALSE
)

spatial_pall <- RunUMAP(spatial_pall,
  dims = 1:30,
  reduction = reductionInt, reduction.name = "umap.integrated.rpca"
) %>%
  FindNeighbors(dims = 1:30, reduction = reductionInt)

resolutions <- c(0.05, 0.1, 0.15, 0.2)
for (res in resolutions) {
  cluster_name <- paste0("SCTbySlice_snn_pall_res.", res)

  spatial_pall <- FindClusters(spatial_pall, resolution = res, cluster.name = cluster_name)

  SpatialDimPlot(spatial_pall,
    group.by = cluster_name,
    ncol = 2, pt.size.factor = 150, image.alpha = 0.2
  )
  ggsave(
    filename = paste0(Figures_dir, "/04_Pall/pallClean_res", gsub("\\.", "", as.character(res)), ".tiff"),
    bg = "white", width = 10, height = 15
  )
}

saveRDS(spatial_pall, file = paste0(RDSfiles, "/spatialPallium_integrated.rds"))
spatial_pall <- readRDS(file = paste0(RDSfiles, "/spatialPallium_integrated.rds"))

Idents(spatial_pall) <- "SCTbySlice_snn_pall_res.0.1"
for (clust in unique(Idents(spatial_pall))) {
  spatial_pall <- FindSubCluster(spatial_pall,
    graph.name = "SCTbySlice_snn",
    cluster = clust, resolution = 0.05, subcluster.name = paste0("sub", clust)
  )
  Idents(spatial_pall) <- paste0("sub", clust)
}

SpatialDimPlot(spatial_pall, ncol = 2, pt.size.factor = 150, image.alpha = 0.2)


#-- pick which res to use
Idents(spatial_pall) <- "SCTbySlice_snn_pall_res.0.1"
subFig_dir <- paste0(Figures_dir, "/integratedRes01/")

for (i in levels(spatial_pall@active.ident)) {
  print(i)
  SpatialDimPlot(spatial_pall,
    cells.highlight = WhichCells(spatial_pall, idents = i),
    ncol = 2, pt.size.factor = 150, image.alpha = 0.1
  )
  ggsave(
    filename = sprintf("%sclustering_subSlides_pallRes02_cluster%s.tiff", subFig_dir, i),
    bg = "white", width = 10, height = 15
  )
}


# seperate more regions
#--arco
Idents(spatial_pall) <- "SCTbySlice_snn_pall_res.0.1"
spatial_pall <- FindSubCluster(spatial_pall,
  graph.name = "SCTbySlice_snn",
  cluster = 3, resolution = 0.05, subcluster.name = "sub_Arco"
)
SpatialDimPlot(spatial_pall,
  group.by = "sub_Arco",
  ncol = 2, pt.size.factor = 150, image.alpha = 0.2
)

Idents(spatial_pall) <- "sub_Arco"
subFig_dir <- paste0(Figures_dir, "/sub_Arco/")
for (i in levels(spatial_pall@active.ident)) {
  print(i)
  SpatialDimPlot(spatial_pall,
    cells.highlight = WhichCells(spatial_pall, idents = i),
    ncol = 2, pt.size.factor = 150, image.alpha = 0.2
  )
  ggsave(
    filename = sprintf("%sclustering_subSlides_pallRes02_cluster%s.tiff", subFig_dir, i),
    bg = "white", width = 10, height = 15
  )
}

#--nido/lman/hyper
Idents(spatial_pall) <- "sub_Arco"
spatial_pall <- FindSubCluster(spatial_pall,
  graph.name = "SCTbySlice_snn",
  cluster = 0, resolution = 0.2, subcluster.name = "sub_ArcoNido"
)

Idents(spatial_pall) <- "sub_ArcoNido"
subFig_dir <- paste0(Figures_dir, "/sub_ArcoNido/")
for (i in levels(spatial_pall@active.ident)) {
  print(i)
  SpatialDimPlot(spatial_pall,
    cells.highlight = WhichCells(spatial_pall, idents = i),
    ncol = 2, pt.size.factor = 150, image.alpha = 0.2
  )
  ggsave(
    filename = sprintf("%sclustering_subSlides_pallRes02_cluster%s.tiff", subFig_dir, i),
    bg = "white", width = 10, height = 15
  )
}


#- use this res
SpatialDimPlot(spatial_pall,
  group.by = "sub_Arco",
  ncol = 2, pt.size.factor = 150, image.alpha = 0.2
)
ggsave(
  filename = paste0(Figures_dir, "/resUse.tiff"),
  bg = "white", width = 10, height = 15
)
saveRDS(spatial_pall, file = paste0(RDSfiles, "/spatialPallium_integrated_subRes.rds"))
```

RA seems to be clustering with mylin cluster...
Maybe not do integration and instead just regress variables..?

Best way seems to be subclustering the RA cluster
- Increasing res in all leads to non-informative, such as not pulling RA
- Subclustering Anterior Nido, also doesnt seem informative
--> sticking to Res 0.1 with cluster 3(RA) subclustered



#===== LOAD THE RESULTS FROM ABOVE
```{r}
spatial_pall <- readRDS(file = paste0(RDSfiles, "/spatialPallium_integrated_subRes.rds"))
SpatialDimPlot(spatial_pall,
  group.by = "sub_Arco",
  ncol = 2, pt.size.factor = 150, image.alpha = 0.2
)
```
