The script documents normalization, integration, clustering, and visualization steps used to generate the final analyses
  when using all of the spatial beads.
This reproduces spatial transcriptomics figures from the manuscript using a preprocessed Seurat object.


###===== SET UP
# LOAD LIBRARIES
```{r}
library(Seurat)
library(tidyverse)
library(readxl)
library(patchwork)

source("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/Functions/commonViz.R")

options(future.globals.maxSize = 100000 * 1024^10)
```

# SET UP DIRECTORIES
```{r}
setwd("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_Spatial")
root_dir <- "/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_Spatial"

RDSfiles <- paste0(root_dir, "/RDS_files")
dir.create(RDSfiles, showWarnings = F)

Figures_dir <- paste0(root_dir, "/Figures")
dir.create(Figures_dir, showWarnings = F)

Tables_dir <- paste0(root_dir, "/Tables")
dir.create(Tables_dir, showWarnings = F)
```


# LOAD DATA
```{r}
spatial_all <-
  readRDS(
    file = "/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_Spatial/ForRepo/spatialAll_seurat.rds"
  )

```


###===== INTEGRATE W/ALL TELE
```{r}
DefaultAssay(spatial_all) <- "Spatial"
spatial_all[["Spatial"]] <- split(spatial_all[["Spatial"]], f = spatial_all$Slice)
spatial_all <- SCTransform(spatial_all,
  assay = "Spatial",
  vars.to.regress = c("nCount_Spatial", "nFeature_Spatial"),
  return.only.var.genes = F,
  new.assay.name = "SCTbySlice"
)

reductionUse <- "pca_sctSlice"
spatial_all <- RunPCA(spatial_all, assay = "SCTbySlice", reduction.name = reductionUse)

reductionInt <- paste0("integrated.rpca_", reductionUse)
spatial_all <- IntegrateLayers(
  object = spatial_all, method = RPCAIntegration,
  orig.reduction = reductionUse, new.reduction = reductionInt,
  normalization.method = "SCT", verbose = FALSE
)

spatial_all <- RunUMAP(spatial_all,
  dims = 1:30,
  reduction = "integrated.rpca_pca_sctSlice", reduction.name = "umap.integrated.rpca"
) %>%
  FindNeighbors(dims = 1:30, reduction = "integrated.rpca_pca_sctSlice")

resolutions <- c(0.05, 0.1, 0.15, 0.2)
for (res in resolutions) {
  cluster_name <- paste0("SCTbySlice_snn_pall_res.", res)

  spatial_all <- FindClusters(spatial_all, resolution = res, cluster.name = cluster_name)

  SpatialDimPlot(spatial_all,
    group.by = cluster_name, label = T,
    ncol = 2, pt.size.factor = 150, image.alpha = 0.2
  )
  ggsave(
    filename = paste0(Figures_dir, "/02_allTeleIntgr_subCluster/Tele_res", as.character(res), ".tiff"),
    bg = "white", width = 12, height = 15
  )
}

Idents(spatial_all) <- "SCTbySlice_snn_pall_res.0.05"
for (clust in unique(Idents(spatial_all))) {
  spatial_all <- FindSubCluster(spatial_all,
    graph.name = "SCTbySlice_snn",
    cluster = clust, resolution = 0.025, subcluster.name = paste0("sub", clust)
  )
  Idents(spatial_all) <- paste0("sub", clust)
}
SpatialDimPlot(spatial_all, ncol = 2, pt.size.factor = 150, image.alpha = 0.2)

for (i in levels(spatial_all@active.ident)) {
  print(i)
  SpatialDimPlot(spatial_all,
    cells.highlight = WhichCells(spatial_all, idents = i),
    ncol = 2, pt.size.factor = 150, image.alpha = 0.2
  )
  ggsave(
    filename = sprintf("%s/02_allTeleIntgr_subCluster/res05_subCluster025/allTele_cluster_%s.tiff", Figures_dir, i),
    bg = "white", width = 10, height = 15
  )
}

saveRDS(spatial_all, file = paste0(RDSfiles, "/spatialAll_integrated_20250207.rds"))
```



#===== IF ABOVE ALREADY RAN ====================================================
```{r}
spatial_all <- readRDS(file = paste0(RDSfiles, "/spatialAll_integrated_20250207.rds"))
spatial_all <- JoinLayers(spatial_all, assay = "Spatial")
```


#- Visualize w/all brain
```{r}
spatial_all$palliumBeads <- ifelse(is.na(spatial_all$palliumBeads), "nonPall", "Pall")

spatial_all$Pallium <- ifelse(spatial_all$sub5 %in% c("0_0", "2_2", "3_0", "4_0", "4_1"), "yes", "no")
spatial_all$Pallium <- ifelse(spatial_all$sub5 %in% c("0_1", "1_0"), "both", spatial_all$Pallium)
spatial_all$sub5_split <- ifelse(spatial_all$sub5 %in% c("0_1", "1_0"), paste(spatial_all$sub5, spatial_all$palliumBeads), spatial_all$sub5)

DimPlot(spatial_all, group.by = "sub5_split")
SpatialDimPlot(spatial_all, group.by = "sub5_split", ncol = 2, pt.size.factor = 150, image.alpha = 0.2)

metaColor <- FetchData(spatial_all, vars = c("Pallium", "sub5_split")) %>%
  unique() %>%
  arrange(Pallium)
metaColor$cols <- c("#a5a082", "#c5a082", "#a5a000", "#d4ac0d", colorRampPalette(c("#bdbdbd", "#212121"))(6), colorRampPalette(c("#1a5276", "#aed6f1"))(5))

color_mapping <- setNames(metaColor$cols, metaColor$sub5_split)

DimPlot(spatial_all, group.by = "sub5_split", cols = color_mapping) + coord_equal()

SpatialDimPlot(spatial_all,
  group.by = "sub5_split", ncol = 2, pt.size.factor = 150, image.alpha = 0.2,
  cols = color_mapping
)
ggsave(
  filename = paste0(Figures_dir, "/03_FigsAll/all_subClusters_Cols.svg"),
  bg = "white", width = 10, height = 15
)

SpatialDimPlot(spatial_all,
  group.by = "sub5_split", ncol = 2, pt.size.factor = 150, image.alpha = 0.2,
  label = T, label.size = 3,
  cols = color_mapping
)
ggsave(
  filename = paste0(Figures_dir, "/03_FigsAll/all_subClusters_sub5split_Cols_wLabel.svg"),
  bg = "white", width = 10, height = 15
)

Idents(spatial_all) <- "sub5_split"
for (i in levels(spatial_all@active.ident)) {
  print(i)
  SpatialDimPlot(spatial_all,
    cells.highlight = WhichCells(spatial_all, idents = i),
    ncol = 2, pt.size.factor = 150, image.alpha = 0.2
  )
  ggsave(
    filename = sprintf("%s/02_allTeleIntgr_subCluster/sub5_split/allTele_cluster_%s.tiff", Figures_dir, i),
    bg = "white", width = 10, height = 15
  )
}

u1 <- DimPlot(spatial_all, group.by = "sub5_split", shuffle = T, cols = color_mapping) +
  NoLegend() +
  coord_equal() +
  ggtitle("Cluster") +
  NoAxes()

ggsave(paste0(Figures_dir, "/03_FigsAll/umap_subClusters_Cols.svg"),
  plot = u1,
  bg = "transparent", width = 8, height = 7
)
```

# Genes to plot for for fig 1
```{r}
#- Make Fig 1 Plot
PlotList <- list()
genes4Fig <- c(
  "TCF7L2", "FOXG1", "DRD1",
  "NR2F2", "TBR1",
  "ETV1", "DACH2", "SATB2",
  "PVALB"
)
sliceUse <- "slice13"
colorUse <- "magma"

for (geneUse in genes4Fig) {
  PlotList[[geneUse]] <- SpatialFeaturePlot(spatial_all,
    features = geneUse,
    images = sliceUse,
    ncol = 2, pt.size.factor = 150, image.alpha = 0.2
  ) +
    scale_fill_viridis_c(option = colorUse) + NoLegend() + ggtitle(geneUse)
}

wrap_plots(PlotList, ncol = 3) & theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
ggsave(
  filename = sprintf("%s/10_PaperFigs/MarkersFeatMap_%s_%s.svg", Figures_dir, sliceUse, colorUse),
  bg = "white", width = 5, height = 3.5
)
```
-- from tosches
ZBTB20, SATB1, DACH2, and ETV1 (ER81)

-- chicken paper
including the known DVR-specific transcription factor gene NR2F2 
SATB2 has been reported to represent a mesopallial marker (33)

-- Evolutionary convergence of sensory circuits in the pallium of amniotes
TBR1: pallium marker
FOXG1: telencephalon marker

#-- Picking pallium clusters
```{r}
#- adding pallium info
spatial_all <- AddMetaData(spatial_all, metadata = beadsInPall)
SpatialDimPlot(spatial_all, group.by = "palliumBeads", ncol = 2, pt.size.factor = 150, image.alpha = 0.2)
spatial_all$cluster_pall <- paste0(spatial_all$sub5, spatial_all$palliumBeads)

Idents(spatial_all) <- "cluster_pall"
for (i in levels(spatial_all@active.ident)) {
  print(i)
  SpatialDimPlot(spatial_all,
    cells.highlight = WhichCells(spatial_all, idents = i),
    ncol = 2, pt.size.factor = 150, image.alpha = 0.2
  )
  ggsave(
    filename = sprintf("%s/PalliumNCluster/cluster%s.tiff", Figures_dir, i),
    bg = "white", width = 10, height = 15
  )
}
```


#-- AVG EXPRESSION PLOT
# cluster naming
```{r}
spatial_all$sub5_split_name <- case_when(
  spatial_all$sub5_split == "0_0" ~ "Nido&Hyper",
  spatial_all$sub5_split == "1_0 Pall" ~ "PallBorders-1",
  spatial_all$sub5_split == "0_1 Pall" ~ "PallBorders-2",
  spatial_all$sub5_split == "2_0" ~ "SubPall",
  spatial_all$sub5_split == "2_1" ~ "AreaX",
  spatial_all$sub5_split == "2_2" ~ "HVC",
  spatial_all$sub5_split == "3_0" ~ "Meso",
  spatial_all$sub5_split == "4_0" ~ "Arco",
  spatial_all$sub5_split == "4_1" ~ "RA",
  spatial_all$sub5_split == "1_0 nonPall" ~ "NonPall-1",
  spatial_all$sub5_split == "6_0" ~ "NonPall-2",
  spatial_all$sub5_split == "1_1" ~ "NonPall-3",
  spatial_all$sub5_split == "1_2" ~ "Thal-1",
  spatial_all$sub5_split == "0_1 nonPall" ~ "Thal-2",
  spatial_all$sub5_split == "5_0" ~ "Thal-3",
)
```


# avg expression
```{r}
genesPlot <- c(genes4Fig, "MBP")
genesPlot <- genesPlot[!grepl("TCF7L2", genesPlot)]
genesPlot <- factor(genesPlot, levels = genesPlot)

clustAvg <- AverageExpression(
  object = spatial_all, assays = "SCTbySlice",
  features = genesPlot, layer = "scale.data", group.by = "sub5_split_name",
  return.seurat = T
)


clustersPlot <- c(
  "SubPall", "AreaX",
  "Arco", "RA", "Nido&Hyper", "HVC", "Meso",
  "PallBorders-1", "PallBorders-2"
)

htMap <- avgExpr_heatmap(clustAvg, genesPlot, clustersUse = "sub5_split_name", colOpt = "magma", scaleOpt = T)
htMap2 <- htMap + ylim(clustersPlot)

ggsave(
  filename = paste0(Figures_dir, "/10_PaperFigs/Tele_AvgExpr.svg"), plot = htMap2,
  width = 3.5, height = 3, bg = "white"
)
```

#-- for figure 1 and supp2
```{r}
toJoin <- FetchData(spatial_all, vars = c("sub5_split", "sub5_split_name")) %>% unique()
metaColor_2 <- left_join(metaColor, toJoin)
color_mapping <- setNames(metaColor_2$cols, metaColor_2$sub5_split_name)
write_csv(metaColor_2, file = paste0(Tables_dir, "/paperTables/spatialClusterNamingAndColors.csv"))

DimPlot(spatial_all,
  group.by = "sub5_split_name", cols = color_mapping,
  label = T, repel = T, label.size = 5, label.color = "#1eb4aa"
) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/10_PaperFigs/umap_telenClusters.svg"),
  bg = "white", width = 8, height = 5
)

#-- to show clusters across all slices
# spatial_all_JIC <- spatial_all
spatial_all@images <- spatial_all@images[c("slice24", "slice13", "slice23", "slice12", "slice21", "slice11", "slice20")]

sptDimPlt <- SpatialDimPlot(spatial_all,
  group.by = "sub5_split_name", ncol = 7, pt.size.factor = 150, image.alpha = 0.2,
  cols = color_mapping
) & NoLegend() & scale_x_reverse()
ggsave(
  filename = paste0(Figures_dir, "/10_PaperFigs/allTeleNslices_subClusters_Cols.svg"), plot = sptDimPlt,
  bg = "white", width = 20, height = 3
)

#-- show region marker genes across all slices
genes4SuppFig <- c(
  "FOXG1", "TCF7L2",
  "DRD1",
  "NR2F2", "TBR1",
  "ETV1", "DACH2", "SATB2",
  "PVALB", "MBP"
)

PlotList_supp <- list()
for (geneUse in genes4SuppFig) {
  PlotList_supp[[geneUse]] <- SpatialFeaturePlot(spatial_all,
    features = geneUse,
    ncol = 7, pt.size.factor = 150, image.alpha = 0.2
  ) &
    scale_fill_viridis_c(option = "magma") & NoLegend() & ggtitle(geneUse) & scale_x_reverse()
}

PlotList_supp_plots <- wrap_plots(PlotList_supp, ncol = 1) & theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
ggsave(
  filename = sprintf("%s/10_PaperFigs/Supp/MarkersFeatMap_AllTeleNSlices.svg", Figures_dir), plot = PlotList_supp_plots,
  bg = "white", width = 20, height = 3 * length(genes4SuppFig)
)
```


```{r}
saveRDS(spatial_all, file = paste0(RDSfiles, "/02_script_output/spatialAll_seurat_20260716.rds"))
```

