---
title: "Making the figures for the paper - Fig. 2"
output: html_notebook
---

This figure will focus on the astrocytes

###========== SET UP
# LOAD LIBRARIES
```{r}
library(Seurat)
library(SeuratDisk)
library(tidyverse)
library(speckle)
library(ggpubr)
library(patchwork)
library(readxl)
library(scCustomize)

source("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/Functions/cell_proportionPlots_20250128.R")
source("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/Functions/CommonLevelsColors.R")
source("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/Functions/stackplots.R")

options(future.globals.maxSize = 100000 * 1024^10)
options(timeout = 100000)

```


# SET UP DIRECTORIES
```{r}
setwd('/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_paper')
root_dir <- '/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_paper'

RDSfiles <- paste0(root_dir, "/RDS_files")
dir.create(RDSfiles, showWarnings = F)

Figures_dir <- paste0(root_dir, "/Figures")
dir.create(Figures_dir, showWarnings = F)

Tables_dir <- paste0(root_dir, "/Tables")
dir.create(Tables_dir, showWarnings = F)

```

# LOAD DATA AND PREP
```{r}
seuObj <- readRDS("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/06_Taxonomy/AllenClustering/RDS_files/OutputUse/seuObj_iterCluster_clean_20250131.rds")

seuObj$RegionAcronym <- factor(seuObj$RegionAcronym, levels = lvl_order)

```

# MODIFY META DATA AS NEEDED
```{r}
unique(seuObj$Allen_classLvl2)
seuObj$Allen_classLvl0 <- ifelse(seuObj$Allen_classLvl2 %in% c("Glutamatergic", "GABAergic"),
                                 "Neurons",
                                 "NonNeurons")

# seuObj$cellClass <- ifelse(seuObj$cellClass_lvl2 %in% c("GABAergic", "OPC_Oligo"),
#                                as.character(seuObj$cellClass_lvl2),
#                                as.character(seuObj$cellClass_lvl3))
# DimPlot(seuObj, group.by = "cellClass")
```

# COMMON
```{r}
metaToPlot <- metaToPlot <- FetchData(seuObj, vars = c("orig.ident", "RegionAcronym", "Allen_classLvl0", "cellClass_lvl2", "cellClass_lvl3", "cellClass"))
metaToPlot$RegionAcronym <- factor(metaToPlot$RegionAcronym, levels = lvl_order)

song_vs_nonsong <- list(c("HVC", "NCM"), c("HVC", "NC"), c("LMAN", "An"), c("RA", "Arco"))

ratioStacks <- 5
heightUse <- 4
widthUse <- 8


```



###--- Fig components

# snRNAseq UMAPS + STACKPLOTS
```{r}
#-- NonNeuN / NeuN
custom_colors <- c("#7a9699", "#aa9f96")

# u1_neuN <- DimPlot(seuObj, group.by = "Allen_classLvl0", cols = custom_colors) +
#   coord_equal() + theme_void() + NoLegend() + theme(plot.title = element_blank())

s1_neuN <- stackedbarplot_colors(metaToPlot,"RegionAcronym", "Allen_classLvl0", horizontal = T) + NoLegend() + coord_fixed(ratioStacks)

#-- Glia subclass / NeuN
# custom_colors <- c("#855C75FF", "#D9AF6BFF", "#AF6458FF", "#736F4CFF", "#526A83FF", "#625377FF", "#68855CFF", "#9C9C5EFF", "#A06177FF", "#8C785DFF", "#467378FF", "#7C7C7CFF")
custom_colors <- c("#768E6D", "#5C798B", "#A38C78", "#A1768F", "#D8B56D", "#8C8C8C")

u2_subGlia <- DimPlot(seuObj, group.by = "cellClass_lvl2", 
                      cols = custom_colors) + 
  coord_equal() + theme_void() + theme(plot.title = element_blank()) + NoLegend()
s2_subGlia <- stackedbarplot_colors(metaToPlot, "RegionAcronym", "cellClass_lvl2", horizontal = T) +
  NoLegend() + coord_fixed(ratioStacks)


ggsave(filename = paste0(Figures_dir, "/Fig2/umaps.tiff"), plot = u2_subGlia,
       bg = NA, width = 5, height = 5)
ggsave(filename = paste0(Figures_dir, "/Fig2/umaps.svg"), plot = u2_subGlia,
       bg = "white", width = 5, height = 5)


s2_subGlia_noLeg <- s2_subGlia +
theme(axis.text.y = element_blank(),  
      axis.ticks.y = element_blank(),
      axis.title.y = element_blank())

panB <- s1_neuN / s2_subGlia
panB

ggsave(filename = paste0(Figures_dir, "/Fig2/stacks.tiff"), plot = panB,
     bg = NA, width = 7, height = 7)
ggsave(filename = paste0(Figures_dir, "/Fig2/stacks.svg"), plot = panB,
     bg = "transparent", width = 7, height = 7)


#--expression inset
sox9_exp <- FeaturePlot_scCustom(seuObj, features = "SOX9")+ coord_equal() + theme_void()

ggsave(filename = paste0(Figures_dir, "/Fig2/featPlot_sox9.tiff"), plot = sox9_exp,
   bg = NA, width = 2, height = 2)
ggsave(filename = paste0(Figures_dir, "/Fig2/featPlot_sox9.svg"), plot = sox9_exp,
   bg = "white", width = 2, height = 2)

```

##=== Barplots
# scRNAseq NonNeuN : NeuN
```{r}
cell_ratios <- table(seuObj[[c("cellClass_lvl3",  "orig.ident", "RegionAcronym")]]) %>% 
  as.data.frame() %>% 
  filter(Freq > 0) %>% 
  group_by(orig.ident) %>%
  mutate(
    Num_freq = sum(Freq[cellClass_lvl3 %in% c("Astro", "Epen", "OPC", "TOL", "Oligo", "Endo", "Micro")]),
    Denom_freq = sum(Freq[cellClass_lvl3 %in% c("MGE_Pvalb", "MGE_Sst", "CGE_Lamp5", "CGE_Vip", "CGE_Vip_Sncg", "LGE", "Glutamatergic", "Neurogenic")])
  ) %>% 
  mutate(Ratio = Num_freq / Denom_freq) %>% 
  select(orig.ident, RegionAcronym, Ratio) %>% 
  unique()
cell_ratios_glia <- cell_ratios

plot_list <- list()
for (pair in song_vs_nonsong) {
  cell_ratios_use <- cell_ratios %>% 
    filter(RegionAcronym %in% c(pair))
  cell_ratios_use$RegionAcronym <- factor(cell_ratios_use$RegionAcronym, levels = pair)

  plot_list[[paste(pair, collapse = " and ")]] <- 
    ggboxplot(cell_ratios_use, x = "RegionAcronym", y = "Ratio",
              color = "RegionAcronym", palette = "uchicago",
              add = "jitter") + 
    stat_compare_means(aes(label = ifelse(..p.signif.. != "ns", ..p.signif..,  paste0("p=", ..p.format..))), 
                         label.x = 1.5,
                         size = 4,
                         fontface = "bold") +
    labs(x = NULL, y = "Glia : Neurons") +
    scale_y_continuous(
      limits = c(0, NA),
      breaks = seq(0, 10, by = 0.1)
      ) +
    NoLegend() +
    theme(
      axis.line = element_line(color = "#060606"),
      axis.ticks = element_line(color = "#060606")
      )
}

bpGlia <- wrap_plots(plot_list, ncol = length(plot_list)) +
  plot_layout(axis_titles = "collect")

bpGlia

```


# scRNAseq Astro : NeuN
```{r}
cell_ratios <- table(seuObj[[c("cellClass_lvl3", "orig.ident", "RegionAcronym")]]) %>% 
  as.data.frame() %>% 
  filter(Freq > 0) %>% 
  group_by(orig.ident) %>%
  mutate(
    Num_freq = sum(Freq[cellClass_lvl3 %in% c("Astro")]),
    Denom_freq = sum(Freq[cellClass_lvl3 %in% c("MGE_Pvalb", "MGE_Sst", "CGE_Lamp5", "CGE_Vip", "CGE_Vip_Sncg", "LGE", "Glutamatergic", "Neurogenic")])
  ) %>% 
  mutate(Ratio = Num_freq / Denom_freq) %>% 
  select(orig.ident, RegionAcronym, Ratio) %>% 
  unique()
cell_ratios_astro <- cell_ratios

plot_list <- list()
for (pair in song_vs_nonsong) {
  cell_ratios_use <- cell_ratios %>% 
    filter(RegionAcronym %in% c(pair))
  cell_ratios_use$RegionAcronym <- factor(cell_ratios_use$RegionAcronym, levels = pair)
  

  plot_list[[paste(pair, collapse = " and ")]] <- 
    ggboxplot(cell_ratios_use, x = "RegionAcronym", y = "Ratio",
              color = "RegionAcronym", palette = "uchicago",
              add = "jitter") + 
    stat_compare_means(aes(label = ifelse(..p.signif.. != "ns", ..p.signif..,  paste0("p=", ..p.format..))), 
                       label.x = 1.5,
                       size = 4,
                       fontface = "bold") +
    labs(x = NULL, y = "Astro : Neurons") +
    scale_y_continuous(
      limits = c(0, NA),
      breaks = seq(0, 10, by = 0.1)
      ) +
    NoLegend()   +
    theme(
      axis.line = element_line(color = "#060606"),
      axis.ticks = element_line(color = "#060606")
      )
}

bpAstro <- wrap_plots(plot_list, ncol = length(plot_list)) +
  plot_layout(axis_titles = "collect")

bpAstro

```

Levels: Astro Epen OPC TOL Oligo Endo Micro MGE_Pvalb MGE_Sst CGE_Lamp5 CGE_Vip CGE_Vip_Sncg LGE Neurogenic Glutamatergic

# stackplots together
```{r}
barplots_comb <- wrap_plots(bpGlia, bpAstro, ncol = 1)

ggsave(filename = paste0(Figures_dir, "/Fig2/scBarplots.tiff"), plot = barplots_comb,
       bg = NA, width = widthUse/1.4, height = heightUse/1.2)
ggsave(filename = paste0(Figures_dir, "/Fig2/scBarplots.svg"), plot = barplots_comb,
   bg = "transparent", width = widthUse/1.4, height = heightUse/1.2)

```


###===== Proportions to Chicken
# Load chk dataset
```{r}
ChkRDSfiles <- '/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/13_chicken/RDS_files'
Gg_adult <- readRDS(file = paste0(ChkRDSfiles, "/ChkZaremba25_wOrthos_sct.rds"))

```

# Make proportion calculations
```{r}
chk_meta <- Gg_adult@meta.data
# DimPlot(Gg_adult, group.by = "tg_class", label = T) + coord_equal() + NoLegend()

chk_cell_ratios <- chk_meta %>% 
  group_by(individual, dissection) %>%
  mutate(
    Astro_freq = sum(tg_class == "Astrocytes"),
    Glia_freq = sum(tg_class %in% c("Astrocytes", "Immune cells", "Oligodendrocyte lineage", "Vascular")),
    Neurons_freq = sum(tg_class %in% c("Excitatory neurons", "Inhibitory neurons"))
  ) %>% 
  mutate(astro_ratio = Astro_freq / Neurons_freq) %>% 
  mutate(glia_ratio = Glia_freq / Neurons_freq) %>% 
  select(individual, dissection, astro_ratio, glia_ratio) %>% 
  unique() %>% 
  rename(orig.ident = individual, RegionAcronym = dissection) 
chk_cell_ratios$group <- "chick"

```


# Join w/ZF
```{r}
cell_ratios_astro <- rename(cell_ratios_astro, astro_ratio = Ratio)
cell_ratios_glia <- rename(cell_ratios_glia, glia_ratio = Ratio)

ratiosSong <- full_join(cell_ratios_astro, cell_ratios_glia)
ratiosSong$group <- ifelse(ratiosSong$RegionAcronym %in% c("HVC", "LMAN", "RA"), "song", "nonSong") 
  
zf_chk_ratios <- full_join(ratiosSong, chk_cell_ratios)

```

# Make plot
```{r}
my_comparisons <- list( c("song", "nonSong"), c("song", "chick"), c("nonSong", "chick"))
zf_chk_ratios$group <- factor(zf_chk_ratios$group, levels = c("song", "nonSong", "chick"))

zfNchk_glia <- ggboxplot(zf_chk_ratios, x = "group", y = "glia_ratio",
          color = "group", palette = "uchicago",
          add = "jitter") + NoLegend() +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif")+ # Add pairwise comparisons p-value
  stat_compare_means(label.y = 1.5) +
  theme(
    axis.line = element_line(color = "#060606"),
    axis.ticks = element_line(color = "#060606")
  ) +
  scale_y_continuous(
limits = c(0, 2))

zfNchk_astro <- ggboxplot(zf_chk_ratios, x = "group", y = "astro_ratio",
          color = "group", palette = "uchicago",
          add = "jitter") + NoLegend() +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif")+ # Add pairwise comparisons p-value
  stat_compare_means(label.y = 1) +
  theme(
    axis.line = element_line(color = "#060606"),
    axis.ticks = element_line(color = "#060606")
    ) +
  scale_y_continuous(
    limits = c(0, 1))

#-save
barplots_wChk <- wrap_plots(zfNchk_glia, zfNchk_astro, ncol = 1) +
  plot_layout(axis_titles = "collect")

ggsave(filename = paste0(Figures_dir, "/Fig2/scBarplots_chkRatios.tiff"), plot = barplots_wChk,
       bg = NA, width = 2, height = 4)
ggsave(filename = paste0(Figures_dir, "/Fig2/scBarplots_chkRatios.svg"), plot = barplots_wChk,
   bg = "transparent", width = 2, height = 4)

```





###======= BARPLOTS FOR IHC VALIDATIONS
```{r}
IHC <- read_excel(paste0(Tables_dir, "/RawIHC_Sox9NeuNDapi_counts_byJudith.xlsx"), 
                  sheet = "ConfocalImages") %>%
       unique() %>%
       drop_na()

IHC$compare <- paste(IHC$Bird, IHC$Hemisphere, sep = "_")

IHC <- IHC %>% 
    mutate(GliaNeuN = `Dapi+NeuN-` / NeuN)%>% 
    mutate(AstroNeuN = `Sox9+NeuN-` / NeuN) %>%
  group_by(Bird, Hemisphere, Region, Slide, Slice) %>%
    mutate(AvgGliaNeuN = mean(GliaNeuN)) %>% 
    mutate(AvgAstroNeuN = mean(AstroNeuN)) %>% 
    mutate(AvgGlia = mean(`Dapi+NeuN-`)) %>% 
    mutate(AvgAstro = mean(`Sox9+NeuN-`)) %>% 
    mutate(AvgNeuN = mean(NeuN)) %>% 
  group_by(Bird, Hemisphere, Region) %>% 
    mutate(AvgBirdHemGlia = mean(AvgGliaNeuN)) %>% 
    mutate(AvgBirdHemAstro = mean(AvgAstroNeuN)) %>% 
    mutate(AvgBirdHemGlia = mean(AvgGlia)) %>% 
    mutate(AvgBirdHemAstro = mean(AvgAstro)) %>% 
    mutate(AvgBirdHemNeuN = mean(AvgNeuN))

write_csv(IHC, file = paste0(Tables_dir, "/FigS2/IHC_quantifications.csv"))

boxPlotsLists <- list()
regions <- c("HVC", "LMAN", "RA")
for (rgn in regions) {
  
  #- filter
  data_filtered <- IHC %>%
    filter(Region %in% c(rgn, paste0("Outside", rgn))) %>% 
  select(Bird, Region, Hemisphere, AvgBirdHemGlia, AvgBirdHemAstro, compare) %>% 
    unique()
  
  #- plot
  for (plotThis in c("AvgBirdHemGlia", "AvgBirdHemAstro")) {
        boxPlotsLists[[paste(rgn, plotThis, sep = "_")]] <- 
          ggpaired(data_filtered, x = "Region", y = plotThis,
                   id = "compare",
            color = "Region", palette = "uchicago",
            line.color = "grey") +
            stat_compare_means(paired = TRUE,
                               aes(label = ifelse(..p.signif.. != "ns", ..p.signif..,  paste0("p=", ..p.format..))), 
                               label.x = 1.5,
                               size = 4,
                               fontface = "bold") +
            labs(x = NULL, y = plotThis) +
            scale_y_continuous(
              limits = c(0, NA),
              breaks = seq(0, 10, by = 0.5)
              ) +
          NoLegend()+
            theme(
              axis.line = element_line(color = "#060606"),
              axis.ticks = element_line(color = "#060606")
            )
        }
  }

gliaBox <- wrap_plots(boxPlotsLists[grepl("Glia", names(boxPlotsLists))]) +
  plot_layout(axis_titles = "collect")

astroBox <- wrap_plots(boxPlotsLists[!grepl("Glia", names(boxPlotsLists))]) +
  plot_layout(axis_titles = "collect")

IHCbox <- gliaBox / astroBox
IHCbox

ggsave(filename = paste0(Figures_dir, "/Fig2/ihcBarplots.tiff"), plot = IHCbox,
       bg = NA, width = widthUse/1.8, height = heightUse/1.3)
ggsave(filename = paste0(Figures_dir, "/Fig2/ihcBarplots.svg"), plot = IHCbox,
       bg = "transparent", width = widthUse/1.8, height = heightUse/1.3)

```


######======================= Tables =================================##########

```{r}
meta4props <- FetchData(seuObj, vars = c("RegionAcronym", "Allen_classLvl0"))

meta4props_percent <- meta4props %>%
  count(RegionAcronym, Allen_classLvl0) %>%        
  group_by(RegionAcronym) %>%                   
  mutate(percent = round(n / sum(n) * 100, digits = 1)) %>%       
  select(-n) %>%
  pivot_wider(
    names_from  = RegionAcronym,
    values_from = percent,
    values_fill = 0
  )

write_csv(meta4props_percent, file = paste0(Tables_dir, "/SuppTable_pct_GliaByRegion.csv"))

```






######===================== Supplementary figures ======================########


#-- DEGs of astros across regions
# Makers
```{r}
astro_seu <- subset(seuObj, subset = clusterName3 == "Astro(12)")
Idents(astro_seu) <- "RegionAcronym"

astro_degs_wilcox <- FindAllMarkers(astro_seu,
                               assay = "SCTorigIdent3",
                               logfc.threshold = 0.1,
                               min.pct = 0.05, 
                               min.diff.pct = 0.15,
                               only.pos = T,
                               test.use = "wilcox",
                               recorrect_umi = F
                               )

astro_degs_wilcox <- astro_degs_wilcox %>% filter(p_val_adj < 0.05)
write_csv(astro_degs_wilcox, file = paste0(Tables_dir, "/FigS2/astroDEGS_byRegion_wlcx_log0.1_minPct0.05_pctDiff0.15.csv"))

```
The roc method only finds 5 genes, not surprising giving the lack of difference..
Plotted all the genes from wilcox but not very informative 
Wilcox w/ logFC@1 aand minPct@0.1, finds 264 genes (significant), so will decrease threshold

W/default threshold, now 8,340 genes --> 6k significant, now getting the top

# Visualize
```{r}
astro_seu <- ScaleData(astro_seu, features = astro_degs_wilcox$gene)

astroMrks_heat <- DoHeatmap(subset(astro_seu, downsample = 700), 
          features = astro_degs_wilcox$gene,
          group.colors = custom_colors_rgn2,
          size = 3)

ggsave(filename = paste0(Figures_dir, "/FigS2/heatPlot_astroMrks.tiff"), plot = astroMrks_heat,
       bg = "white", width = 7, height = 12)

DotPlot_scCustom(astro_seu, features = unique(astro_degs_wilcox$gene), flip_axes = T) 

```




#-- IHC raw/density of cells
```{r}
IHC <- read_csv(file = paste0(Tables_dir, "/FigS2/IHC_quantifications.csv"))

boxPlotsLists <- list()
regions <- c("HVC", "LMAN", "RA")
for (rgn in regions) {
  
  #- filter
  data_filtered <- IHC %>%
    filter(Region %in% c(rgn, paste0("Outside", rgn))) %>% 
  select(Bird, Region, Hemisphere, AvgBirdHemGlia, AvgBirdHemAstro, AvgBirdHemNeuN, compare) %>% 
    unique()
  
  #- plot
  for (plotThis in c("AvgBirdHemGlia", "AvgBirdHemAstro", "AvgBirdHemNeuN")) {
        boxPlotsLists[[paste(rgn, plotThis, sep = "_")]] <- 
          ggpaired(data_filtered, x = "Region", y = plotThis,
                   id = "compare",
            color = "Region", palette = "uchicago",
            line.color = "grey") +
            stat_compare_means(paired = TRUE,
                               aes(label = ifelse(..p.signif.. != "ns", ..p.signif..,  paste0("p=", ..p.format..))), 
                               label.x = 1.5,
                               size = 4,
                               fontface = "bold") +
            labs(x = NULL, y = plotThis) +
            scale_y_continuous(
              limits = c(0, 12),
              breaks = seq(0, 12, by = 2)
              ) +
          NoLegend()+
            theme(
              axis.line = element_line(color = "#060606"),
              axis.ticks = element_line(color = "#060606")
            )
        }
  }

gliaBox_dense <- wrap_plots(boxPlotsLists[grepl("Glia", names(boxPlotsLists))]) +
  plot_layout(axis_titles = "collect")

astroBox_dense <- wrap_plots(boxPlotsLists[grepl("Astro", names(boxPlotsLists))]) +
  plot_layout(axis_titles = "collect")

neuronBox_dense <- wrap_plots(boxPlotsLists[grepl("NeuN", names(boxPlotsLists))]) +
  plot_layout(axis_titles = "collect")

IHCbox <- gliaBox_dense / astroBox_dense / neuronBox_dense
IHCbox

ggsave(filename = paste0(Figures_dir, "/FigS2/ihcBarplots_rawNumbers.svg"), plot = IHCbox,
       bg = "transparent", width = 4, height = 6)

```





