###========== SET UP
# LOAD LIBRARIES
```{r}
library(Seurat)
library(tidyverse)
library(patchwork)
library(readr)
library(SeuratWrappers)
library(SeuratData)
library(scCustomize)
library(Azimuth)
library(pheatmap)

source("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/Functions.R")
source("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/Functions/cell_proportionPlots_20240903.R")

options(future.globals.maxSize = 100000 * 1024^10)
```

# SET UP DIRECTORIES
```{r}
setwd("/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/05c_ClassifyCells_subcluster_083024")
root_dir <- "/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/05c_ClassifyCells_subcluster_083024"

RDSfiles <- paste0(root_dir, "/RDS_files")
dir.create(RDSfiles, showWarnings = F)

Figures_dir <- paste0(root_dir, "/Figures")
dir.create(Figures_dir, showWarnings = F)

Tables_dir <- paste0(root_dir, "/Tables")
dir.create(Figures_dir, showWarnings = F)
```

# LISTS
```{r}
CanMarkers <- c(
  "SLC17A6", # glut cells,
  # "NR2E1", #adult NSCs
  # "SOX4", #neuronally committed precursors
  # "DCX", #migrating neuroblasts
  "GAD1", # gaba cells
  "PDGFRA", # OPCs
  "MBP", # Oligos
  "LRIG1", # Astro
  "CSF1R", # Microglia
  "FLT1" # Endothelial
)
```


#=== LOAD DATA AND PREP
```{r}
seuObj <- readRDS(file = "/project/Neuroinformatics_Core/Roberts_lab/s433904/PalliumEvo_AnalysisR/04_CheckDbl_otherQC/RDS_files/IntegratedRPCA_CxGclass_20240826.rds")

seuObj_2 <- subset(seuObj, subset = integrated.rpca_clusters %in% c("46", "53", "50", "51"), invert = T)

# cell class
seuObj_2$integrated.rpca_CellClass <- "Empty"
cluster_class_mapping <- list(
  Glutamatergic = c(
    "2", "5", "20", "35", "22", "11", "32", "24", "27", "3", "47", "14", "0", "9", "30", "19", "37", "21", "29", "15", "26",
    "39", "43", "16"
  ),
  GABAergic = c(
    "52", "34", "1", "42", "33", "17", "13", "23", "36", "10", "38", "28",
    "41"
  ),
  OPC_Oligo = c("12", "49", "45", "7"),
  Endo = c("18", "40", "54"), # endo
  Astro = c("25", "6", "8", "44", "48", "4"),
  Microglia = c("31")
)

for (class_name in names(cluster_class_mapping)) {
  cluster_names <- cluster_class_mapping[[class_name]]
  cluster_cells <- seuObj_2@active.ident %in% cluster_names
  seuObj_2$integrated.rpca_CellClass[cluster_cells] <- class_name
}

DimPlot(seuObj_2, reduction = "umap.integrated.rpca", group.by = "integrated.rpca_CellClass") + coord_equal()
```




###==== Re-integrate with all cells by orig.ident

1) sct and pca by each orig.ident 
2) Subset by cell class (need to reclassify)
3) Don't re-normalize, only run new graph and clusters and umap. 

# Integrate by orig.ident 
```{r}
# prep
DefaultAssay(seuObj_2) <- "RNA"
seuObj_2 <- DietSeurat(seuObj_2, assays = "RNA")

seuObj_2 <- JoinLayers(seuObj_2)
seuObj_2[["RNA"]] <- split(seuObj_2[["RNA"]], f = seuObj_2$orig.ident)

# std process each w/just PCA
assayUse <- "SCTorigIdent"
reductionUse <- paste0("pca.", assayUse)

seuObj_2 <- SCTransform(seuObj_2, new.assay.name = assayUse, return.only.var.genes = T)
seuObj_2 <- RunPCA(seuObj_2, assay = assayUse, reduction.name = reductionUse)

reductionInt <- paste0("integrated.rpca_", reductionUse)
seuObj_2 <- IntegrateLayers(
  object = seuObj_2, method = RPCAIntegration,
  orig.reduction = reductionUse, new.reduction = reductionInt,
  normalization.method = "SCT", verbose = FALSE
)
saveRDS(seuObj_2, file = paste0(RDSfiles, "/Integrated_sctByOrig_20240830.rds"))

seuObj_2 <- FindNeighbors(seuObj_2,
  dims = 1:30, reduction = reductionInt,
  graph.name = c("rpca_nn", "rpca_snn")
) %>%
  FindClusters(resolution = 1, graph.name = paste0("rpca", "_snn"), cluster.name = paste0(reductionInt, "_clusters_", "res1"))
seuObj_2 <- RunUMAP(seuObj_2,
  dims = 1:30,
  reduction = reductionInt,
  reduction.name = paste0("umap.", reductionInt)
)
saveRDS(seuObj_2, file = paste0(RDSfiles, "/Integrated_sctByOrig_umap_20240830.rds"))

seuObj_2 <- readRDS(file = paste0(RDSfiles, "/Integrated_sctByOrig_umap_20240830.rds")) #---- LAST SAVED !!!!!!!!!!!!

# visualize
DimPlot(seuObj_2, group.by = "orig.ident", cols = custom_colors, reduction = paste0("umap.", reductionInt)) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/umap_byOrigIdent.tiff"),
  bg = "white", width = 8, height = 7
)

seuObj_2 <- BuildClusterTree(seuObj_2, features = CanMarkers, reorder = T)
seuObj_2$integrated.rpca_pca.SCTorigIdent_clusters_res1 <- Idents(seuObj_2)

DotPlot(seuObj_2, features = CanMarkers) + RotatedAxis()
ggsave(
  filename = paste0(Figures_dir, "/dotplot_CanMarkers.tiff"),
  bg = "white", width = 5, height = 9
)

stackMeta <- FetchData(seuObj_2, vars = c("orig.ident", "integrated.rpca_pca.SCTorigIdent_clusters_res1"))
stackedbarplot_colors(stackMeta, "integrated.rpca_pca.SCTorigIdent_clusters_res1", "orig.ident")
ggsave(
  filename = paste0(Figures_dir, "/stackplot_OrigIdent.tiff"),
  bg = "white", width = 10, height = 15
)

#-- relcassify cell classes
# cell class
seuObj_2$integrated.rpca_CellClass_2 <- "Empty"
cluster_class_mapping <- list(
  Glutamatergic = c(
    "18", "3", "7", "9", "19", "23", "38", "14", "24", "28", "2", "0", "6", "33", "25", "32", "16", "22",
    "17"
  ),
  GABAergic = c(
    "8", "29", "36", "10", "27", "31", "34", "41", "15", "20", "40", "13", "35",
    "37"
  ),
  OPC_Oligo = c(
    "12",
    "4", "39"
  ),
  Endo = c("21"),
  Astro = c("26", "1", "5", "11"),
  Microglia = c("30")
)

for (class_name in names(cluster_class_mapping)) {
  cluster_names <- cluster_class_mapping[[class_name]]
  cluster_cells <- seuObj_2@active.ident %in% cluster_names
  seuObj_2$integrated.rpca_CellClass_2[cluster_cells] <- class_name
}

DimPlot(seuObj_2, group.by = "integrated.rpca_CellClass_2") + coord_equal()
DimPlot(seuObj_2, group.by = "orig.ident", cols = custom_colors) + coord_equal()
DimPlot(seuObj_2, label = T) + coord_equal()

IDs <- unique(seuObj_2$integrated.rpca_pca.SCTorigIdent_clusters_res1)
Idents(seuObj_2) <- "integrated.rpca_pca.SCTorigIdent_clusters_res1"
plot_list <- list()
for (id in IDs) {
  p <- DimPlot(seuObj_2,
    reduction = paste0("umap.", reductionInt),
    cells.highlight = CellsByIdentities(seuObj_2, idents = id), order = id
  ) + coord_equal()
  plot_list[[as.character(id)]] <- p
}
combined_plot <- wrap_plots(plot_list, ncol = 10)
ggsave(
  filename = paste0(Figures_dir, "/umap_HiglightByRPCA.tiff"), plot = combined_plot,
  bg = "white", width = 40, height = 25
)
```


###==== REMOVE CLUSTERS ONCE AGAIN and reprocess
```{r}
FeaturePlot(seuObj_2, features = "nFeature_RNA", max.cutoff = 2000)
VlnPlot_scCustom(
  seurat_object = seuObj_2,
  features = c("nFeature_RNA", "nFeature_SCTorigIdent"),
  group.by = "integrated.rpca_CellClass_2", pt.size = 0.1
)

seuObj_3 <- subset(seuObj_2, subset = integrated.rpca_pca.SCTorigIdent_clusters_res1 %in% c("33", "35"), invert = T)

# prep
DefaultAssay(seuObj_3) <- "RNA"
seuObj_3 <- DietSeurat(seuObj_3, assays = "RNA")

seuObj_3 <- JoinLayers(seuObj_3)
seuObj_3[["RNA"]] <- split(seuObj_3[["RNA"]], f = seuObj_3$orig.ident)

# std process each w/just PCA
assayUse <- "SCTorigIdent2"
reductionUse <- paste0("pca.", assayUse)

seuObj_3 <- SCTransform(seuObj_3, new.assay.name = assayUse, return.only.var.genes = T)
seuObj_3 <- RunPCA(seuObj_3, assay = assayUse, reduction.name = reductionUse)

reductionInt <- paste0("integrated.rpca_", reductionUse)
seuObj_3 <- IntegrateLayers(
  object = seuObj_3, method = RPCAIntegration,
  orig.reduction = reductionUse, new.reduction = reductionInt,
  normalization.method = "SCT", verbose = FALSE
)
saveRDS(seuObj_3, file = paste0(RDSfiles, "/Integrated_sctByOrig_20240830.rds"))

seuObj_3 <- FindNeighbors(seuObj_3,
  dims = 1:30, reduction = reductionInt,
  graph.name = c("rpca_nn", "rpca_snn")
) %>%
  FindClusters(resolution = 1, graph.name = paste0("rpca", "_snn"), cluster.name = paste0(reductionInt, "_clusters_", "res1"))
seuObj_3 <- RunUMAP(seuObj_3,
  dims = 1:30,
  reduction = reductionInt,
  reduction.name = paste0("umap.", reductionInt)
)

# saveRDS(seuObj_3, file = paste0(RDSfiles, "/Integrated_sctByOrig_umap_20240904.rds"))
# seuObj_3 <- readRDS(file = paste0(RDSfiles, "/Integrated_sctByOrig_umap_20240904.rds"))

# visualize
DimPlot(seuObj_3, reduction = paste0("umap.", reductionInt), label = T) + coord_equal()
DimPlot(seuObj_3, group.by = "orig.ident", cols = custom_colors, reduction = paste0("umap.", reductionInt)) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/umap_byOrigIdent.tiff"),
  bg = "white", width = 8, height = 7
)

seuObj_3 <- BuildClusterTree(seuObj_3, features = CanMarkers, reorder = T)
seuObj_3$integrated.rpca_pca.SCTorigIdent2_clusters_res1 <- Idents(seuObj_3)

DotPlot(seuObj_3, features = CanMarkers) + RotatedAxis()
ggsave(
  filename = paste0(Figures_dir, "/dotplot_CanMarkers.tiff"),
  bg = "white", width = 5, height = 9
)

stackMeta <- FetchData(seuObj_3, vars = c("orig.ident", "integrated.rpca_pca.SCTorigIdent2_clusters_res1", "integrated.rpca_CellClass_2"))
stackedbarplot_colors(stackMeta, "integrated.rpca_pca.SCTorigIdent2_clusters_res1", "orig.ident")
ggsave(
  filename = paste0(Figures_dir, "/stackplot_origIdent.tiff"),
  bg = "white", width = 10, height = 15
)
stackedbarplot(stackMeta, "integrated.rpca_pca.SCTorigIdent2_clusters_res1", "integrated.rpca_CellClass_2")
ggsave(
  filename = paste0(Figures_dir, "/stackplot_CellClass2.tiff"),
  bg = "white", width = 20, height = 15
)
# cell class doesn't seem to work so need to reclassify

IDs <- unique(seuObj_3$integrated.rpca_pca.SCTorigIdent2_clusters_res1)
Idents(seuObj_3) <- "integrated.rpca_pca.SCTorigIdent2_clusters_res1"
plot_list <- list()
for (id in IDs) {
  p <- DimPlot(seuObj_3,
    reduction = paste0("umap.", reductionInt),
    cells.highlight = CellsByIdentities(seuObj_3, idents = id), order = id
  ) + coord_equal()
  plot_list[[as.character(id)]] <- p
}
combined_plot <- wrap_plots(plot_list, ncol = 10)
ggsave(
  filename = paste0(Figures_dir, "/umap_HiglightByRPCA.tiff"), plot = combined_plot,
  bg = "white", width = 40, height = 25
)
```

# reclassify cells again
```{r}
unique(seuObj_3$integrated.rpca_pca.SCTorigIdent2_clusters_res1)

seuObj_3$integrated.rpca_CellClass_3 <- "Unassigned"
cluster_class_mapping <- list(
  Glutamatergic = c(
    "9", "21", "23", "37", "14", "27", "22", "13", "4", "17", "2", "0", "5", "28", "16", "19", "30", "33",
    "18"
  ),
  GABAergic = c(
    "8", "29", "35", "40", "26", "32", "34", "38", "15", "24", "10", "11",
    "36"
  ),
  OPC_Oligo = c(
    "12",
    "6", "39"
  ),
  Endo = c("20"),
  Astro = c("25", "1", "3", "7"),
  Microglia = c("31")
)

# Loop through each class in the cluster_class_mapping and assign the cell class based on cluster identities
for (class_name in names(cluster_class_mapping)) {
  clusters <- cluster_class_mapping[[class_name]]
  seuObj_3$integrated.rpca_CellClass_3[seuObj_3$integrated.rpca_pca.SCTorigIdent2_clusters_res1 %in% clusters] <- class_name
}

DimPlot(seuObj_3, label = T) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/umap_clusters.tiff"),
  bg = "white", width = 8, height = 7
)

DimPlot(seuObj_3, group.by = "integrated.rpca_CellClass_3") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/umap_CellClass3.tiff"),
  bg = "white", width = 8, height = 7
)

FeaturePlot(seuObj_3, features = "nFeature_SCTorigIdent2", coord.fixed = T)


### --- removing 40 and re-running UMAP >:3
seuObj_4 <- subset(seuObj_3, subset = integrated.rpca_pca.SCTorigIdent2_clusters_res1 == "40", invert = T)
seuObj_4 <- RunUMAP(seuObj_4,
  dims = 1:30,
  reduction = reductionInt,
  reduction.name = paste0("umap.", reductionInt)
)

# fixing classification
seuObj_4$integrated.rpca_CellClass_3[seuObj_4$integrated.rpca_pca.SCTorigIdent2_clusters_res1 %in% "36"] <- "GABAergic"

DimPlot(seuObj_4, group.by = "integrated.rpca_CellClass_3") + coord_equal()
DimPlot(seuObj_4, label = T) + coord_equal()
DimPlot(seuObj_4, group.by = "orig.ident", cols = custom_colors) + coord_equal()
```

# RE-NORMALIZE AND STD PROCESS TO CHECK W/OUT CLUSTER 40
Will continue with this one and reclassify cell one more.
This is excluding cluster 40, which looked like ad GABA-micro doublets
```{r}
# prep
DefaultAssay(seuObj_4) <- "RNA"
seuObj_4 <- DietSeurat(seuObj_4, assays = "RNA")

seuObj_4 <- JoinLayers(seuObj_4)
seuObj_4[["RNA"]] <- split(seuObj_4[["RNA"]], f = seuObj_4$orig.ident)

# std process each w/just PCA
assayUse <- "SCTorigIdent3"
reductionUse <- paste0("pca.", assayUse)

seuObj_4 <- SCTransform(seuObj_4, new.assay.name = assayUse, return.only.var.genes = T)
seuObj_4 <- RunPCA(seuObj_4, assay = assayUse, reduction.name = reductionUse)

reductionInt <- paste0("integrated.rpca_", reductionUse)
seuObj_4 <- IntegrateLayers(
  object = seuObj_4, method = RPCAIntegration,
  orig.reduction = reductionUse, new.reduction = reductionInt,
  normalization.method = "SCT", verbose = FALSE
)
saveRDS(seuObj_4, file = paste0(RDSfiles, "/Integrated_sctByOrig_20240905.rds"))

seuObj_4 <- FindNeighbors(seuObj_4,
  dims = 1:30, reduction = reductionInt,
  graph.name = c("rpca_nn", "rpca_snn")
) %>%
  FindClusters(resolution = 1, graph.name = paste0("rpca", "_snn"), cluster.name = paste0(reductionInt, "_clusters_", "res1"))
seuObj_4 <- RunUMAP(seuObj_4,
  dims = 1:30,
  reduction = reductionInt,
  reduction.name = paste0("umap.", reductionInt)
)

saveRDS(seuObj_4, file = paste0(RDSfiles, "/Integrated_sctByOrig_umap_20240905.rds"))
seuObj_4 <- readRDS(file = paste0(RDSfiles, "/Integrated_sctByOrig_umap_20240905.rds"))

# VISUALIZE
DimPlot(seuObj_4, reduction = paste0("umap.", reductionInt), label = T) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/umap_clusters.tiff"),
  bg = "white", width = 8, height = 7
)

DimPlot(seuObj_4, group.by = "orig.ident", cols = custom_colors, reduction = paste0("umap.", reductionInt)) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/umap_byOrigIdent.tiff"),
  bg = "white", width = 8, height = 7
)

seuObj_4 <- BuildClusterTree(seuObj_4, features = CanMarkers, reorder = T)
seuObj_4$integrated.rpca_pca.SCTorigIdent3_clusters_res1 <- Idents(seuObj_4)

DotPlot(seuObj_4, features = CanMarkers) + RotatedAxis()
ggsave(
  filename = paste0(Figures_dir, "/dotplot_CanMarkers.tiff"),
  bg = "white", width = 5, height = 9
)

stackMeta <- FetchData(seuObj_4, vars = c("orig.ident", "integrated.rpca_pca.SCTorigIdent3_clusters_res1", "integrated.rpca_CellClass_3"))
stackedbarplot_colors(stackMeta, "integrated.rpca_pca.SCTorigIdent3_clusters_res1", "orig.ident")
ggsave(
  filename = paste0(Figures_dir, "/stackplot_origIdent.tiff"),
  bg = "white", width = 10, height = 15
)
stackedbarplot(stackMeta, "integrated.rpca_pca.SCTorigIdent3_clusters_res1", "integrated.rpca_CellClass_3")
ggsave(
  filename = paste0(Figures_dir, "/stackplot_CellClass3.tiff"),
  bg = "white", width = 20, height = 15
)

IDs <- unique(seuObj_4$integrated.rpca_pca.SCTorigIdent3_clusters_res1)
Idents(seuObj_4) <- "integrated.rpca_pca.SCTorigIdent3_clusters_res1"
plot_list <- list()
for (id in IDs) {
  p <- DimPlot(seuObj_4,
    reduction = paste0("umap.", reductionInt),
    cells.highlight = CellsByIdentities(seuObj_4, idents = id), order = id
  ) + coord_equal()
  plot_list[[as.character(id)]] <- p
}
combined_plot <- wrap_plots(plot_list, ncol = 10)
ggsave(
  filename = paste0(Figures_dir, "/umap_HiglightByRPCAclusters.tiff"), plot = combined_plot,
  bg = "white", width = 40, height = 25
)
```
Didn't recover more microglia, which is what I was trying. 
So instead will proceed with current cleaning at subcluster level, and then recluster.
ACTUALLY, will classify micro-endo as one, so might as well reclassify again.

# reclassify cells--v4
```{r}
unique(seuObj_4$integrated.rpca_pca.SCTorigIdent3_clusters_res1)

seuObj_4$integrated.rpca_CellClass_4 <- "Unassigned"
cluster_class_mapping <- list(
  Glutamatergic = c(
    "11", "18", "21", "35", "4", "22", "2", "0", "7", "20", "3", "16",
    "15", "26", "13", "30", "32"
  ),
  GABAergic = c(
    "8", "27", "33", "25", "29", "17", "24", "31", "36", "10", "14", "37",
    "34"
  ),
  OPC_Oligo = c(
    "12",
    "5", "38"
  ),
  Endo_Micro = c(
    "19",
    "28"
  ),
  Astro = c("23", "1", "6", "9")
)

# Loop through each class in the cluster_class_mapping and assign the cell class based on cluster identities
for (class_name in names(cluster_class_mapping)) {
  clusters <- cluster_class_mapping[[class_name]]
  seuObj_4$integrated.rpca_CellClass_4[seuObj_4$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% clusters] <- class_name
}

DimPlot(seuObj_4, group.by = "integrated.rpca_CellClass_4") + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/umap_CellClass4.tiff"),
  bg = "white", width = 8, height = 7
)
```



###===== subclustering by cell Class
modified later on 9/6/24 to make Endo_Micro as one group
```{r}
# seuObj_4 <- readRDS(file = paste0(RDSfiles, "/Integrated_sctByOrig_umap_20240905.rds"))
# seuObj_4$integrated.rpca_CellClass_3[seuObj_4$integrated.rpca_CellClass_3 %in% c("Endo", "Microglia")] <- "Endo_Micro"

CellClasses <- unique(seuObj_4$integrated.rpca_CellClass_4)
# CellClasses <- CellClasses[5] # if want to test

for (cellClass in CellClasses) {
  sub_Seu <- subset(seuObj_4, subset = integrated.rpca_CellClass_3 == cellClass)

  sub_Seu <- FindNeighbors(sub_Seu,
    dims = 1:30, reduction = reductionInt,
    graph.name = c(paste0(cellClass, ".rpca_nn"), paste0(cellClass, ".rpca_snn"))
  ) %>%
    FindClusters(
      resolution = 1, graph.name = paste0(cellClass, ".rpca", "_snn"),
      cluster.name = paste0(reductionInt, "_clusters_", cellClass, "_res1")
    )
  sub_Seu <- RunUMAP(sub_Seu,
    dims = 1:30,
    reduction = reductionInt,
    reduction.name = paste0("umap.", cellClass, ".", reductionInt)
  )

  saveRDS(sub_Seu, paste0(RDSfiles, "/sub", cellClass, "_umap.rds"))

  # figures
  subFig_dir <- paste0(Figures_dir, "/", cellClass, "/")
  dir.create(subFig_dir)

  DimPlot(sub_Seu, reduction = paste0("umap.", cellClass, ".", reductionInt), group.by = "orig.ident", cols = custom_colors, shuffle = T) + coord_equal()
  ggsave(
    filename = paste0(subFig_dir, "umap_byOrigIdent.tiff"),
    bg = "white", width = 8, height = 7
  )

  stackMeta <- FetchData(sub_Seu, vars = c("orig.ident", paste0(reductionInt, "_clusters_", cellClass, "_res1")))
  stackedbarplot_colors(stackMeta, paste0(reductionInt, "_clusters_", cellClass, "_res1"), "orig.ident")
  ggsave(
    filename = paste0(subFig_dir, "stackplot_byOrigIdent.tiff"),
    bg = "white", width = 20, height = 15
  )

  DotPlot(sub_Seu, features = CanMarkers) + RotatedAxis()
  ggsave(
    filename = paste0(subFig_dir, "dotplot_CanMarkers.tiff"),
    bg = "white", width = 5, height = length(unique(sub_Seu@active.ident)) - (length(unique(sub_Seu@active.ident)) * 0.4)
  )

  DimPlot(sub_Seu, reduction = paste0("umap.", cellClass, ".", reductionInt), shuffle = T, label = T) + coord_equal()
  ggsave(
    filename = paste0(subFig_dir, "umap_bySubCluster.tiff"),
    bg = "white", width = 8, height = 7
  )
}
```


#=== check specific subclusters
# Azimuth
```{r}
#--- use azimuth to help classify cells classes quicker
DefaultAssay(seuObj_4) <- "RNA"
seuObj_4 <- JoinLayers(seuObj_4)
seuObj_4 <- RunAzimuth(seuObj_4, reference = "mousecortexref")
metaData <- seuObj_4[[49:57]]

saveRDS(metaData, file = paste0(RDSfiles, "/meta_Predicted_MsCtx.rds"))

DimPlot(seuObj_4, reduction = "umap.integrated.rpca_pca.SCTorigIdent3", group.by = "predicted.cross_species_cluster") + coord_equal()

# make a cluster map, where each predicted cell class is
IDs <- unique(seuObj_4$predicted.subclass)
Idents(seuObj_4) <- "predicted.subclass"
plot_list <- list()
for (id in IDs) {
  p <- DimPlot(seuObj_4,
    reduction = "umap.integrated.rpca_pca.SCTorigIdent3",
    cells.highlight = CellsByIdentities(seuObj_4, idents = id), order = id
  ) + coord_equal()
  plot_list[[as.character(id)]] <- p
}
combined_plot <- wrap_plots(plot_list, ncol = 11)
ggsave(
  filename = paste0(Figures_dir, "/umap_HiglightByPredicSubClass.tiff"), plot = combined_plot,
  bg = "white", width = 40, height = 20
)


metaData <- readRDS(file = paste0(RDSfiles, "/meta_Predicted_MsCtx.rds"))
```


# check specific subclusters for further figures
```{r}
# load specific subclusters and add metaData
subEndoMicro <- readRDS(paste0(RDSfiles, "/subEndo_Micro_umap.rds"))

subEndoMicro <- AddMetaData(subEndoMicro, metaData)

IDs <- unique(subEndoMicro$predicted.subclass)
Idents(subEndoMicro) <- "predicted.subclass"
plot_list <- list()
for (id in IDs) {
  p <- DimPlot(subEndoMicro,
    reduction = "umap.Endo_Micro.integrated.rpca_pca.SCTorigIdent3",
    cells.highlight = CellsByIdentities(subEndoMicro, idents = id), order = id
  ) + coord_equal()
  plot_list[[as.character(id)]] <- p
}
combined_plot <- wrap_plots(plot_list, ncol = 5)
ggsave(
  filename = paste0(Figures_dir, "/Endo_Micro/umap_HiglightByPredicSubClass.tiff"), plot = combined_plot,
  bg = "white", width = 35, height = 30
)


DimPlot(subEndoMicro, reduction = "umap.Endo_Micro.integrated.rpca_pca.SCTorigIdent3", group.by = "predicted.subclass")


#-- load in each subcluster in for loop
subclustered <- list.files(path = paste0(RDSfiles, "/subclustered"))

loaded_objects <- list()
for (file in subclustered) {
  obj_name <- gsub("sub|_umap.rds", "", file)
  loaded_objects[[obj_name]] <- readRDS(paste0(RDSfiles, "/subclustered/", file))
}

for (obj_name in names(loaded_objects)) {
  obj <- loaded_objects[[obj_name]]
  obj <- AddMetaData(obj, metaData)

  IDs <- unique(obj$predicted.subclass)
  Idents(obj) <- "predicted.subclass"
  plot_list <- list()
  for (id in IDs) {
    p <- DimPlot(obj,
      reduction = paste0("umap.", obj_name, ".integrated.rpca_pca.SCTorigIdent3"),
      cells.highlight = CellsByIdentities(obj, idents = id), order = id
    ) + coord_equal()
    plot_list[[as.character(id)]] <- p
  }
  combined_plot <- wrap_plots(plot_list, ncol = 5)
  ggsave(
    filename = paste0(Figures_dir, "/", obj_name, "/umap_HighlightByPredicSubClass.tiff"),
    plot = combined_plot, bg = "white", width = 35, height = 30
  )
}
```


# create a dataframe of cells to remove based on subclustering
```{r}
clusters_to_extract <- list(
  Endo_Micro = c(12, 13),
  Astro = c(18),
  GABAergic = c(26, 27, 28),
  Glutamatergic = c(28)
)

filtered_cells <- list()
for (obj_name in names(clusters_to_extract)) {
  obj <- loaded_objects[[obj_name]]
  cluster_col <- paste0("integrated.rpca_pca.SCTorigIdent3_clusters_", obj_name, "_res1")

  if (cluster_col %in% colnames(obj@meta.data)) {
    cluster_data <- obj@meta.data[[cluster_col]]
    clusters <- clusters_to_extract[[obj_name]]
    cell_ids <- rownames(obj@meta.data)[cluster_data %in% clusters]
    filtered_cells[[obj_name]] <- cell_ids
  } else {
    warning(paste("Cluster column", cluster_col, "not found in", obj_name))
  }
}

#-- Now drop from main seurat object and recluster
seuObj_all <- readRDS(file = paste0(RDSfiles, "/Integrated_sctByOrig_cellClass_20240906.rds"))
DimPlot(seuObj_all, cells.highlight = WhichCells(seuObj_all, cells = filtered_cells)) + coord_equal()

all_cells_to_drop <- unlist(filtered_cells)
seuObj_all_2 <- subset(seuObj_all, cells = all_cells_to_drop, invert = T)

DimPlot(seuObj_all_2, label = T) + coord_equal()
```


# subclassify cells now
```{r}
DotPlot(seuObj_all_2, group.by = "integrated.rpca_CellClass_4", features = CanMarkers)

seuObj_all_2$cellClass_lvl2 <- "Unassigned"
cluster_class_mapping <- list(
  Glutamatergic = c(
    "11", "18", "21", "35", "4", "22", "2", "0", "7", "20", "3", "16",
    "15", "26", "13", "30", "32"
  ),
  GABAergic = c(
    "8", "27", "33", "25", "29", "17", "24", "31", "36", "10", "14", "37",
    "34"
  ),
  OPC_Oligo = c(
    "12",
    "38",
    "5"
  ),
  Endo = c("19"),
  Micro = c("28"),
  Astroependymal = c("23", "1", "6", "9")
)

for (class_name in names(cluster_class_mapping)) {
  clusters <- cluster_class_mapping[[class_name]]
  seuObj_all_2$cellClass_lvl2[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% clusters] <- class_name
}

DimPlot(seuObj_all_2, group.by = "cellClass_lvl2") + coord_equal()

### === lvl3
seuObj_all_2$cellClass_lvl3 <- seuObj_all_2$cellClass_lvl2

# Astro subgroups
seuObj_all_2$cellClass_lvl3[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% c(23)] <- "Epen"
seuObj_all_2$cellClass_lvl3[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% c(1, 6, 9)] <- "Astro"

# #-- subclassify oligos
# obj <- loaded_objects[["OPC_Oligo"]]
# DotPlot(obj, features = c("PDGFRA",
#                           "GPR17", #COPs "SOX6"
#                           "TCF7L2", #NFOLS
#                           "MBP"))
#
#
# Idents(seuObj_all_2) <- "integrated.rpca_pca.SCTorigIdent3_clusters_res1"
# seuObj_all_2 <- FindSubCluster(seuObj_all_2, cluster = "38", graph.name = "rpca_snn",
#                                subcluster.name = "integrated.rpca_pca.SCTorigIdent3_clusters_res1_subcluster",
#                                resolution = 0.2)
#
# DimPlot(seuObj_all_2, label = T,
#         group.by = "integrated.rpca_pca.SCTorigIdent3_clusters_res1_subcluster") + coord_equal()
#
# DotPlot(seuObj_all_2,
#         group.by = "integrated.rpca_pca.SCTorigIdent3_clusters_res1_subcluster",
#         features = c("PDGFRA",
#                      "GPR17", #COPs "SOX6"
#                      "TCF7L2", #NFOLS
#                      "MBP"))   #---------------------------------------- EASIER WITH ALL TOGETHER

seuObj_all_2$cellClass_lvl3[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% c(12)] <- "OPC"
seuObj_all_2$cellClass_lvl3[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% c(38)] <- "TOL"
seuObj_all_2$cellClass_lvl3[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% c(5)] <- "Oligo"

DimPlot(seuObj_all_2, group.by = "cellClass_lvl3") + coord_equal()


# #-- subclassify endos
# obj <- loaded_objects[["Endo_Micro"]]
# DotPlot(obj, features = c("FLT1",
#                           "CALD1", "EBF1", "PDGFRB", "ACTA2", "CSPG4", # peris
#                           "BNC2", "CPED1", "MGP" #VLMC
#                           )) + RotatedAxis()


#-- NEUROGENIC CELLS
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

seuObj_all_2 <- CellCycleScoring(
  seuObj_all_2,
  s.features = s.genes,
  g2m.features = g2m.genes,
  assay = "SCTorigIdent3",
  set.ident = F
)

DimPlot(seuObj_all_2, group.by = "Phase") + coord_equal()

subFig_dir <- paste0(Figures_dir, "/", "Neurogenic", "/")
dir.create(subFig_dir)

seuObj_Neurons <- subset(seuObj_all_2, subset = cellClass_lvl3 %in% c("Glutamatergic", "GABAergic"))
ImmMarkers <- c("NR2E1", "SOX4", "NECTIN3", "DCX")
DotPlot(seuObj_Neurons, features = c(ImmMarkers, CanMarkers[1:2])) + RotatedAxis()
ggsave(
  filename = paste0(subFig_dir, "/dotplot_ImmNeuMarkers.tiff"),
  bg = "white", width = 5, height = 9
)

FeaturePlot(seuObj_Neurons, features = c(ImmMarkers), order = T, coord.fixed = T)

seuObj_all_2$cellClass_lvl3[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% c(37, 34, 15, 13)] <- "Neurogenic"
DimPlot(seuObj_all_2, group.by = "cellClass_lvl3") + coord_equal()


#-- GABAergic subtypes
seuObj_all_2@meta.data <- seuObj_all_2@meta.data[, -c(15:45)]
metaData <- readRDS(file = paste0(RDSfiles, "/meta_Predicted_MsCtx.rds"))
seuObj_all_2 <- AddMetaData(seuObj_all_2, metaData)
DimPlot(seuObj_all_2, group.by = "predicted.subclass", label = T) + coord_equal()

IDs <- unique(seuObj_all_2$predicted.subclass)
Idents(seuObj_all_2) <- "predicted.subclass"
plot_list <- list()
for (id in IDs) {
  p <- DimPlot(seuObj_all_2,
    reduction = paste0("umap.integrated.rpca_pca.SCTorigIdent3"),
    cells.highlight = CellsByIdentities(seuObj_all_2, idents = id), order = id
  ) + coord_equal()
  plot_list[[as.character(id)]] <- p
}
combined_plot <- wrap_plots(plot_list, ncol = 5)
ggsave(
  filename = paste0(Figures_dir, "/umap_HighlightByPredicSubClass.tiff"),
  plot = combined_plot, bg = "white", width = 35, height = 30
)

seuObj_all_2$cellClass_lvl3[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% c(8)] <- "MGE_Pvalb"
seuObj_all_2$cellClass_lvl3[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% c(17, 24, 31, 36)] <- "MGE_Sst"
seuObj_all_2$cellClass_lvl3[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% c(27, 33)] <- "CGE_Lamp5"
seuObj_all_2$cellClass_lvl3[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% c(25)] <- "CGE_Vip"
seuObj_all_2$cellClass_lvl3[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% c(29)] <- "CGE_Vip_Sncg"
seuObj_all_2$cellClass_lvl3[seuObj_all_2$integrated.rpca_pca.SCTorigIdent3_clusters_res1 %in% c(10, 14)] <- "LGE"

DimPlot(seuObj_all_2, group.by = "cellClass_lvl3") + coord_equal()
new_order <- c(
  "Astro", "Epen",
  "OPC", "TOL", "Oligo",
  "Endo", "Micro",
  "MGE_Pvalb", "MGE_Sst", "CGE_Lamp5", "CGE_Vip", "CGE_Vip_Sncg", "LGE", # GABAergic
  "Neurogenic",
  "Glutamatergic"
)
seuObj_all_2$cellClass_lvl3 <- factor(seuObj_all_2$cellClass_lvl3, levels = new_order)
DimPlot(seuObj_all_2,
  group.by = "cellClass_lvl3",
  label = T,
  cols = c(
    colorRampPalette(c("#bf8e56", "#f3cca0"))(2),
    colorRampPalette(c("#e0acf3", "#7e3f94"))(3),
    "#ba9d9a", "#923830",
    colorRampPalette(c("#c5d8c8", "#108b24"))(6),
    "#83cacc",
    "#4278c0"
  )
) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/umap_CellClass_lvl3.tiff"),
  bg = "white", width = 8, height = 7
)

saveRDS(seuObj_all_2, file = paste0(RDSfiles, "/Integrated_sctByOrig_cellClassLvl3_20241001.rds"))
seuObj_all_2 <- readRDS(paste0(RDSfiles, "/Integrated_sctByOrig_cellClassLvl3_20241001.rds"))
```

#--- Class levels coloring
```{r}
# add other higher levels
seuObj_all_2$cellClass_lvl0 <- "unassigned"
seuObj_all_2$cellClass_lvl0[seuObj_all_2$cellClass_lvl2 %in% c("Glutamatergic", "GABAergic")] <- "Neurons"
seuObj_all_2$cellClass_lvl0[seuObj_all_2$cellClass_lvl2 %in% c("OPC_Oligo", "Astroependymal", "Micro", "Endo")] <- "Non-neurons"
DimPlot(seuObj_all_2, group.by = "cellClass_lvl0", cols = c("#7f8ea8", "#a79c88")) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/umap_CellClass_lvl0.tiff"),
  bg = "white", width = 8, height = 7
)

seuObj_all_2$cellClass_lvl1a <- "unassigned"
seuObj_all_2$cellClass_lvl1a[seuObj_all_2$cellClass_lvl2 %in% c("Glutamatergic", "GABAergic")] <- "Neurons"
seuObj_all_2$cellClass_lvl1a[seuObj_all_2$cellClass_lvl2 %in% c("OPC_Oligo", "Astroependymal")] <- "Non-neuronal"
seuObj_all_2$cellClass_lvl1a[seuObj_all_2$cellClass_lvl2 %in% c("Micro", "Endo")] <- "Non-neural"
DimPlot(seuObj_all_2, group.by = "cellClass_lvl1a", cols = c("#7f8ea8", "#dac092", "#927c55")) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/umap_CellClass_lvl1a.tiff"),
  bg = "white", width = 8, height = 7
)

seuObj_all_2$cellClass_lvl1b <- "unassigned"
seuObj_all_2$cellClass_lvl1b[seuObj_all_2$cellClass_lvl2 %in% c("Glutamatergic", "GABAergic")] <- "Neurons"
seuObj_all_2$cellClass_lvl1b[seuObj_all_2$cellClass_lvl2 %in% c("OPC_Oligo", "Astroependymal", "Micro")] <- "Glia"
seuObj_all_2$cellClass_lvl1b[seuObj_all_2$cellClass_lvl2 %in% c("Endo")] <- "Endo"
DimPlot(seuObj_all_2, group.by = "cellClass_lvl1b", cols = c("#b8b5b4", "#a89089", "#7f8ea8")) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/umap_CellClass_lvl1b.tiff"),
  bg = "white", width = 8, height = 7
)

seuObj_all_2$cellClass_lvl2 <- factor(seuObj_all_2$cellClass_lvl2, levels = c("Glutamatergic", "GABAergic", "Astroependymal", "OPC_Oligo", "Micro", "Endo"))
DimPlot(seuObj_all_2, group.by = "cellClass_lvl2", cols = c("#4278c0", "#6aa974", "#d3a269", "#9c77a9", "#923830", "#ba9d9a")) + coord_equal()
ggsave(
  filename = paste0(Figures_dir, "/umap_CellClass_lvl2.tiff"),
  bg = "white", width = 8, height = 7
)
```
Higher levels I have decided are inaccurate, but will change later

#-- Remove clusters with high nCount_RNA by subcluster
```{r}
VlnPlot_scCustom(seuObj_all_2,
  features = "nFeature_RNA",
  group.by = "cellClass_lvl0",
  colors_use = c("#7f8ea8", "#a79c88"),
  plot_boxplot = T,
  pt.size = 0
)

meta_nFeatureRNA <- seuObj_all_2[[c("nFeature_RNA", "cellClass_lvl0", "cellClass_lvl2")]]
meta_nFeatureRNA <- meta_nFeatureRNA %>% rownames_to_column(var = "cellID")
metadata_outliers <- meta_nFeatureRNA %>%
  group_by(cellClass_lvl0) %>%
  mutate(
    Q1 = quantile(nFeature_RNA, 0.25),
    Q3 = quantile(nFeature_RNA, 0.75),
    IQR = Q3 - Q1,
    Outlier_byRNAfeat = (nFeature_RNA < (Q1 - 6 * IQR)) | (nFeature_RNA > (Q3 + 6 * IQR))
  ) %>%
  column_to_rownames(var = "cellID")

summary_table <- metadata_outliers %>%
  group_by(cellClass_lvl2) %>%
  summarize(
    Q1 = unique(Q1),
    Q3 = unique(Q3),
    IQR = unique(IQR),
    cutoff_6byIQR = (Q3 + 6 * IQR),
    Outlier_Count = sum(Outlier_byRNAfeat),
    PercentRemoved = (sum(Outlier_byRNAfeat) / n()) * 100
  )
write.csv(summary_table, file = paste0(Tables_dir, "/summaryDrop_subClass_nFeatureRNA_table.csv"), row.names = FALSE)

metadata_outliers_2 <- select(metadata_outliers, Outlier_byRNAfeat)
seuObj_all_2 <- AddMetaData(seuObj_all_2, metadata = metadata_outliers_2)

lvls <- c("cellClass_lvl0", "cellClass_lvl2", "cellClass_lvl3")
for (lvl in lvls) {
  VlnPlot_scCustom(seuObj_all_2,
    features = "nFeature_RNA",
    group.by = lvl,
    split.by = "Outlier_byRNAfeat",
    colors_use = c("#4278c0", "#d3a269")
  )
  ggsave(
    filename = paste0(Figures_dir, "/vlnplot_", lvl, "RNAfeatDrop.tiff"),
    bg = "white", width = 7, height = 8
  )
}

DimPlot(seuObj_all_2, group.by = "Outlier_byRNAfeat") + coord_equal()

saveRDS(seuObj_all_2, file = paste0(RDSfiles, "/Integrated_sctByOrig_cellClassLvl3_20241001.rds"))

seuObj_all_2 <- subset(seuObj_all_2, subset = Outlier_byRNAfeat == "FALSE")
saveRDS(seuObj_all_2, file = paste0(RDSfiles, "/OutputObject_toUse/Integrated_sctByOrig_cellClassLvl3_20241001_notOutliers.rds"))
```

# DOES INCREASING RESOLUTION SPLIT CLUSTER 35
```{r}
seuObj_all_2 <- readRDS(file = paste0(RDSfiles, "/OutputObject_toUse/Integrated_sctByOrig_cellClassLvl3_20241001_notOutliers.rds"))

assayUse <- "SCTorigIdent"
reductionUse <- paste0("pca.", assayUse)
reductionInt <- paste0("integrated.rpca_", reductionUse)

for (resUse in seq(from = 1.2, to = 2.4, by = 0.2)) {
  seuObj_all_2 <- FindClusters(seuObj_all_2,
    resolution = resUse, graph.name = paste0("rpca", "_snn"),
    cluster.name = paste0(reductionInt, "_clusters_", resUse)
  )
}

saveRDS(seuObj_all_2, file = paste0(RDSfiles, "/OutputObject_toUse/Integrated_sctByOrig_cellClassLvl3_notOutliers_res1to24_20241018.rds"))
```

